---
title: "Merging back clean data with Contros corrected data file"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```


```{r packages}
library(tidyverse)
library(lubridate)
library(zoo)
# library(dygraphs)
# library(xts)
```


# CTD and pCO~2~ data

## Merging summarized data sets


```{r load_summarized_data}
# Load Sensor and HydroC data ---------------------------------------------

CTD <- read_csv(here::here("Data/_summarized_data_files",
                              "Tina_V_Sensor_Profiles_Transects.csv"),
                   col_types = list("pCO2" = col_double())) %>%
  rename(pCO2_analog = pCO2)

HC <- read_csv(here::here("Data/_summarized_data_files", "Tina_V_HydroC.csv"))

# Time offset correction ----------------------------------------------

# Time offset was determined by comparing zeroing reads from Sensor and HC
# in the plots produced in the section Time stamp synchronzity below
# before applying this correction

CTD <- CTD %>%
  mutate(day = yday(date_time),
         date_time = if_else(day >= 206 & day <= 220,
                             date_time - 80, date_time - 10)) %>%
  select(-day)

# Merge Sensor and HydroC data --------------------------------------------

df <- full_join(CTD, HC) %>%
  arrange(date_time)

rm(HC, CTD)
```


## Interpolation to common time stamp

CTD and auxillary recordings (15 sec measurment interval) are interpolated to HydroC time stamps (first 10 sec, than 1 sec measurement interval) when gaps between observations are not larger than 20. Thereafter, HydroC readings not falling in regular transects/profilings are removed, by removing rows with NA depth values. Furthermore, CTD readings without corresponding HydroC reading are removed, except during periods when HydroC was not operating.

```{r interpolation}

# Interpolate Sensor data to HydroC timestamp

df <-
  df %>%
  mutate(dep = na.approx(dep, na.rm = FALSE, maxgap = 20),
         sal = na.approx(sal, na.rm = FALSE, maxgap = 20),
         tem = na.approx(tem, na.rm = FALSE, maxgap = 20),
         pCO2_analog = na.approx(pCO2_analog, na.rm = FALSE, maxgap = 20),
         pH = na.approx(pH, na.rm = FALSE, maxgap = 20),
         V_pH = na.approx(V_pH, na.rm = FALSE, maxgap = 20),
         O2 = na.approx(O2, na.rm = FALSE, maxgap = 20),
         Chl = na.approx(Chl, na.rm = FALSE, maxgap = 20)) %>%
  filter(!is.na(dep)) %>% #remove HC readings not falling in regular transects/profilings
  fill(ID, type, station, cast) %>%
  filter(!is.na(deployment) | is.na(pCO2_analog)) # removes CTD readings without corresponding HydroC reading, except during periods when HydroC was not operating

```


```{r merge_back_with_HydroC_raw_data}

HC <-
  read_csv2(here::here("Data/TinaV/Sensor/HydroC-pCO2/corrected_Contros",
                       "parameter&pCO2s(method 43).txt"),
            col_names = c("date_time", "Zero_HC", "Flush_HC", "p_NDIR",
                          "p_in", "T_control", "T_gas", "%rH_gas",
                          "Signal_raw", "Signal_ref", "T_sensor",
                          "pCO2", "Runtime", "nr.ave")) %>%
  mutate(date_time = dmy_hms(date_time))

df <- df %>% 
  select(-c(pCO2))

df <- full_join(df, HC)

df <- df %>% 
  filter(!is.na(ID))

rm(list=setdiff(ls(), "df"))

```


```{r clean_douplicated_time_stamps}

# add counter for date_time observations

df <- df %>%
  add_count(date_time)

# find triplicated time stamp and select only first observation, and merge

df_no_triple <- df %>%
  filter(n <= 2)

df_triple_clean <- df %>%
  filter(n > 2) %>%
  group_by(date_time) %>% 
  slice(1)

df <- full_join(df_no_triple, df_triple_clean)

#no doubles detected

rm(list=setdiff(ls(), "df"))


```


```{r safe_merged_data_file}
write_csv(df, here::here("Data/_merged_data_files", "BloomSail_CTD_HydroC_Contros_clean.csv"))
```



# Tasks / open questions

- for some reason the merged data file (Contros raw + clean df) has more lines than df, bot after left_join of full_join and filter NA's in dep column
