---
title: "MLD"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---


```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r packages}
library(tidyverse)
library(seacarb)
library(oce)
library(marelac)
library(patchwork)
library(metR)
library(lubridate)
```


```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# Approach

In order to test how (and how well) the depth-integrated C~T~ estimates can be reproduced if only surface CO2 data were available, the BloomSail observations were restricted to those made in surface water and following reconstruction approaches were tested:

- Mixed layer depth: Integration of surface observation across the MLD, assuming homogenious vertical patterns
- CT profile reconstruction: Vertical reconstruction of incremental C~T~ changes based on profiles of incremental changes in temperature
- Temperature penetration depth:  Integration of surface observation across the temperture penetration depth, assuming similar vertical extension as for CT drawdown.

Note: The reconstruction of CT profiles and the integration across the temperature penetrations depth should produce very similar results. However, the latter avoids to create misinterpretable information about the vertical distribution of CT.

# Sensor data

1m gridded, downcast profiles were used.

```{r read_CT_cum}

ts_profiles_ID <-
  read_csv(here::here("Data/_merged_data_files", "ts_profiles_ID_long_cum_MLD.csv"))

ts_profiles_ID <- ts_profiles_ID %>% 
  select(ID, date_time_ID, dep, CT = value, CT_diff = value_diff, CT_cum = value_cum, sign, rho_lim, MLD)

ts_profiles_ID <- ts_profiles_ID %>% 
  mutate(rho_lim = as.factor(rho_lim))

```


```{r read_ts_data}

ts_profiles_ID_long_cum <-
  read_csv(here::here("Data/_merged_data_files", "ts_profiles_ID_long_cum.csv"))

ts_profiles_ID_long_cum <- ts_profiles_ID_long_cum %>% 
  filter(var %in% c("CT", "tem")) %>% 
  mutate(ID = as.factor(ID)) %>% 
  select(-date_ID)

```

# Removing c~T~ at depth

CO~2~ data other than 3.5 m water depth were removed.

```{r remove_CT_depth}

ts_profiles_ID <- ts_profiles_ID %>% 
  mutate(CT = if_else(dep == 3.5, CT, NaN),
         rho_lim = as.factor(rho_lim))

```


# MLD approach

MLD calculation was previously described in the [CT dynamics](https://jens-daniel-mueller.github.io/BloomSail/CT_dynamics.html) chapter.

## Timeseries

For comparison to MLD, the effective penetration depth of NCP, z~eff~, was calculated as the ratio of the inremental, depth-integrated change of C~T~, divided by the change in surface C~T~, for all cases where the change in surface C~T~ was negative.

```{r NCP_penetration_depth}

ts_profiles_ID_surface <- ts_profiles_ID %>% 
  filter(sign=="neg",
         rho_lim == "0.5",
         dep == 3.5) %>% 
  group_by(ID, date_time_ID) %>% 
  summarise(CT_diff_surf = mean(CT_diff, na.rm = TRUE)) %>% 
  ungroup()

ts_profiles_ID_i <- ts_profiles_ID %>% 
  filter(sign=="neg",
         rho_lim == "0.5") %>% 
  group_by(ID, date_time_ID) %>% 
  summarise(CT_diff_i = sum(CT_diff, na.rm = TRUE))

zeff <- inner_join(ts_profiles_ID_surface, ts_profiles_ID_i)
rm(ts_profiles_ID_surface, ts_profiles_ID_i)

zeff <- zeff %>% 
  mutate(zeff = CT_diff_i / CT_diff_surf)

```



```{r MLD_time_series, fig.asp=0.4}

ts_profiles_ID %>% 
  ggplot()+
  geom_hline(yintercept = 0)+
  geom_line(data = zeff, aes(date_time_ID, zeff, linetype="zeff"))+
  geom_line(aes(date_time_ID, MLD, col=rho_lim, linetype="MLD"))+
  scale_y_reverse()+
  scale_color_viridis_d(name="Rho limit")+
  scale_linetype(name="Estimate")+
  labs(y="Depth (m)")+
  theme(axis.title.x = element_blank())

rm(zeff)

```

## iC~T~ calculation

Integrated C~T~ depletion was calculated as the product of observed incremental C~T~ changes in surface waters and the respective mixed layer depth. 

```{r NCP_calculation}

ts_profiles_ID_surface <- ts_profiles_ID %>% 
  filter(dep==3.5) %>% 
  group_by(rho_lim) %>% 
  arrange(date_time_ID) %>% 
  mutate(iCT_diff = CT_diff * MLD / 1000,
         iCT_cum = cumsum(replace_na(iCT_diff, 0))) %>% 
  ungroup()
  
```


## Incremental and cumulative timeseries

Total incremental and cumulative C~T~ changes inbetween cruise dates were calculated.


```{r iCT_timeseries_MLD, fig.asp=1}

p_iCT <- ts_profiles_ID_surface %>% 
  ggplot(aes(date_time_ID, iCT_diff, fill= rho_lim))+
  geom_hline(yintercept = 0)+
  geom_col(col="black", position = "dodge")+
  scale_y_continuous(breaks = seq(-100, 100, 0.2))+
  scale_fill_viridis_d()+
  labs(y="integrated CT changes [mol/m2]")+
  theme(axis.title.x = element_blank())

p_iCT_cum <- ts_profiles_ID_surface %>% 
  ggplot(aes(date_time_ID,  iCT_cum, 
             col=rho_lim))+
  geom_line()+
  geom_hline(yintercept = 0)+
  scale_color_viridis_d()+
  scale_y_continuous(breaks = seq(-100, 100, 0.2))+
  theme(strip.background = element_blank(),
        strip.text = element_blank())+
  labs(y="integrated, cumulative CT changes [mol/m2]", x="date")


(p_iCT / p_iCT_cum)+
  plot_layout(guides = 'collect')

rm(p_iCT, p_iCT_cum)

rm(ts_profiles_ID, ts_profiles_ID_surface)

```


# Temperature approach


```{r read_ts_data_tem_appraoch}

ts_profiles_ID <-
  read_csv(here::here("Data/_merged_data_files", "ts_profiles_ID.csv"))

ts_profiles_ID <- ts_profiles_ID %>% 
  mutate(CT = if_else(dep == 3.5, CT, NaN),
         ID = as.factor(ID)) %>% 
  select(-date_ID)

```


## C~T~ vs temperature

As primary production (negative changes in C~T~) and increase in seawater temperature have a common driver (light), the relation between both changes was investigated.

```{r CT_and_temperature_changes_in_surface}

ts_profiles_ID_diff <- ts_profiles_ID %>% 
  drop_na() %>% 
  arrange(date_time_ID) %>%
  mutate(CT_diff = CT - lag(CT, default = first(CT)),
         tem_diff = tem - lag(tem, default = first(tem)),
         factor = CT_diff / tem_diff,
         factor = if_else(is.na(factor), 0, factor))

ts_profiles_ID_diff %>% 
  ggplot(aes(tem_diff, CT_diff))+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  geom_path()+
  geom_point()

```


## Reconstruction of C~T~ dynamics

The ratio of the incremental change of C~T~ with temperature at the seasurface was applied to calculate the C~T~ in other water depth based on the known change in temperature.

```{r reconctruction_CT_incremental_changes_profiles}

ts_profiles_ID_diff <- ts_profiles_ID_diff %>% 
  select(ID, factor)

ts_profiles_ID <- full_join(ts_profiles_ID, ts_profiles_ID_diff)
rm(ts_profiles_ID_diff)

ts_profiles_ID <- ts_profiles_ID %>%
  arrange(date_time_ID) %>%
  group_by(dep) %>% 
  mutate(tem_diff = tem - lag(tem, default = first(tem))) %>% 
  ungroup()

ts_profiles_ID <- ts_profiles_ID %>% 
  mutate(CT_diff = tem_diff * factor) %>% 
  select(-factor)

```

The reconstructed incremental changes are added up to derive cummulative C~T~ changes throughout the water column.

```{r reconctruction_CT_cumulative_changes_profiles}

ts_profiles_ID_long <- ts_profiles_ID %>% 
  select(ID, date_time_ID, dep, tem = tem_diff, CT = CT_diff) %>% 
  pivot_longer(4:5, names_to = "var", values_to = "value_diff") %>% 
  group_by(dep, var) %>% 
  arrange(date_time_ID) %>% 
  mutate(date_time_ID_diff = as.numeric(date_time_ID - lag(date_time_ID)),
         date_time_ID_ref  = date_time_ID - (date_time_ID - lag(date_time_ID))/2,
         value_diff_daily = value_diff / date_time_ID_diff,
         value_cum = cumsum(value_diff)) %>% 
  ungroup()

```

## Profiles of incremental changes

Changes of seawater parameters at each depth were reconstructed from one cruise day to the next and divided by the number of days inbetween.

```{r incremental_changes_profiles}

ts_profiles_ID_long %>% 
  arrange(dep) %>% 
  ggplot(aes(value_diff, dep, col=ID))+
  geom_vline(xintercept = 0)+
  geom_point()+
  geom_path()+
  scale_y_reverse()+
  scale_color_viridis_d()+
  facet_wrap(~var, scales = "free_x")+
  labs(x="Change of value inbetween cruises per day")


```


## Profiles of cumulative changes

Cumulative changes of seawater parameters were calculated at each depth relative to the first cruise day on July 5.

```{r cumulative_changes_profiles}

ts_profiles_ID_long %>% 
  arrange(dep) %>% 
  ggplot(aes(value_cum, dep, col=ID))+
  geom_vline(xintercept = 0)+
  geom_point()+
  geom_path()+
  scale_y_reverse()+
  scale_color_viridis_d()+
  labs(x="Cumulative change of value")+
  facet_wrap(~var, scales = "free_x")

```

## Hovmoeller plots

Hoevmoeller plots were generated for the reconstructed daily and cumulative changes in C~T~. Absolute values are not reproducible with this approach. Furthermore, it meets our expectations 

### Daily changes

```{r plot_hovmoeller_daily, fig.cap="Hovmoeller plots of daily reconstructed changes in C~T~ and temperature. Note: Daily changes are currently plotted against the day when they were observed compared to the previous transect, although plotting against the mean date would be more plausible.", fig.asp=0.5}

bin_CT <- 2.5

ts_profiles_ID_long %>%
  filter(var == "CT") %>% 
  ggplot()+
  geom_contour_fill(aes(x=date_time_ID_ref, y=dep, z=value_diff_daily),
                    breaks = MakeBreaks(bin_CT),
                    col="black")+
  geom_point(aes(x=date_time_ID, y=c(24.5)), size=3, shape=24, fill="white")+
  scale_fill_divergent(breaks = MakeBreaks(bin_CT),
                       guide = "colorstrip",
                       name="CT (µmol/kg)")+
  scale_y_reverse()+
  theme_bw()+
  labs(y="Depth (m)")+
  coord_cartesian(expand = 0)+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

rm(bin_CT)

```

### Cumulative changes

```{r plot_hovmoeller_cum, fig.cap="Hovmoeller plots of daily reconstructed changes in C~T~ and temperature.", fig.asp=0.5}

bin_CT <- 20

ts_profiles_ID_long %>%
  filter(var == "CT") %>% 
  ggplot()+
  geom_contour_fill(aes(x=date_time_ID, y=dep, z=value_cum),
                    breaks = MakeBreaks(bin_CT),
                    col="black")+
  geom_point(aes(x=date_time_ID, y=c(24.5)), size=3, shape=24, fill="white")+
  scale_fill_divergent(breaks = MakeBreaks(bin_CT),
                       guide = "colorstrip",
                       name="CT (µmol/kg)")+
  scale_y_reverse()+
  theme_bw()+
  labs(y="Depth (m)")+
  coord_cartesian(expand = 0)+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())


rm(bin_CT)

```


## iC~T~ time series

Total incremental and cumulative C~T~ changes inbetween cruise dates were calculated for the upper 10 m of the water body.

```{r integrated_NCP}

iCT_10 <- ts_profiles_ID_long %>%
  filter(dep < 10, 
         var == "CT") %>% 
  select(ID, date_time_ID, date_time_ID_ref, CT_diff=value_diff, CT_cum=value_cum) %>% 
  group_by(ID, date_time_ID, date_time_ID_ref) %>% 
  summarise(CT_i_diff = sum(CT_diff)/1000,
            CT_i_cum = sum(CT_cum)/1000) %>% 
  ungroup()

iCT_10 %>% 
  ggplot()+
  #geom_point(data = cruise_dates, aes(date_time_ID, 0), shape=21)+
  geom_col(aes(date_time_ID_ref, CT_i_diff),
           position = "dodge", alpha=0.3)+
  geom_line(aes(date_time_ID, CT_i_cum))+
  scale_color_viridis_d(name="Depth limit (m)")+
  scale_fill_viridis_d(name="Depth limit (m)")+
  labs(y="iCT [mol/m2]", x="")+
  theme_bw()


```



# Heat and C~T~ penetration depth

To investigate the link between temperature rise and CT drawdown, we estimated the mean penetration depth of both parameters, which was defined as the surface change (0-6m) devided by the integrated change across depth. To derive the integrated values, we summed up all negative changes in CT and temperature. For temperature we limited the integration to the depth at which the cummulative signal on July 23 was below 5% of the surface signal.

```{r penetration_depth_tem_integration_depth}

# surface values

diff_surface <- ts_profiles_ID_long_cum %>% 
  filter(dep < 6) %>% 
  group_by(ID, var) %>% 
  summarise(value_diff_surface = mean(value_diff, na.rm = TRUE),
            value_cum_surface  =mean(value_cum, na.rm = TRUE)) %>% 
  ungroup()

ts_profiles_ID_long_cum <- full_join(ts_profiles_ID_long_cum, diff_surface) 
rm(diff_surface)

# directional changes in CT and tem

penetration_depth_CT <- ts_profiles_ID_long_cum %>%
  filter(var == "CT") %>% 
  mutate(value_diff = if_else(value_diff<=0, value_diff, NaN),
         value_cum = if_else(value_cum  <=0, value_cum, NaN))

penetration_depth_tem <- ts_profiles_ID_long_cum %>%
  filter(var == "tem") %>% 
  mutate(value_diff = if_else(value_diff>=0, value_diff, NaN),
         value_cum = if_else(value_cum  >=0, value_cum, NaN))

penetration_depth <- full_join(penetration_depth_CT,
                                  penetration_depth_tem)

# define integration depth

integration_depth_tem <- penetration_depth_tem %>% 
  filter(ID == "180723") %>% 
  mutate(value_cum_rel = 100*value_cum/value_cum_surface)

integration_depth_tem %>% 
  ggplot(aes(value_cum_rel, dep))+
  geom_vline(xintercept = c(5))+
  geom_point()+
  geom_path()+
  scale_y_reverse(name = "Depth (m)")+
  scale_x_continuous(breaks = seq(0,100,10),
                     name = "Cumulative temperature change relative to surface (%)")

integration_depth_tem <- integration_depth_tem %>% 
  filter(value_cum_rel < 5) %>% 
  pull(dep) %>% 
  min()

rm(penetration_depth_CT, penetration_depth_tem)

```

```{r penetration_depth_calculation}


# calculate penetration depths

penetration_depth <- penetration_depth %>% 
  filter(dep < integration_depth_tem) %>% 
  group_by(var, ID, date_time_ID) %>% 
  summarise(value_diff_int = sum(value_diff, na.rm = TRUE),
            value_cum_int = sum(value_cum, na.rm = TRUE),
            value_diff_surface = mean(value_diff_surface, na.rm = TRUE),
            value_cum_surface = mean(value_cum_surface, na.rm = TRUE),
            penetration_depth_diff = value_diff_int/value_diff_surface,
            penetration_depth_cum = value_cum_int/value_cum_surface) %>% 
  ungroup() %>% 
  filter(penetration_depth_diff > 0)

penetration_depth_mean <- penetration_depth %>% 
  group_by(var) %>% 
  summarise(penetration_depth_diff_mean = mean(penetration_depth_diff, na.rm=TRUE),
            penetration_depth_cum_mean = mean(penetration_depth_cum, na.rm=TRUE),
            penetration_depth_diff_sd = sd(penetration_depth_diff, na.rm=TRUE),
            penetration_depth_cum_sd = sd(penetration_depth_cum, na.rm=TRUE)) %>% 
  ungroup()

penetration_depth_mean


```


```{r penetration_depth_CT-tem, fig.asp=1.2}

p_pene_diff <- penetration_depth %>% 
  ggplot(aes(date_time_ID, penetration_depth_diff, col=var))+
  geom_hline(yintercept = 0)+
  geom_hline(data = penetration_depth_mean,
             aes(yintercept = penetration_depth_diff_mean,
                 col=var, linetype = "mean"))+
  geom_line(aes(linetype="cruise"))+
  geom_point()+
  scale_y_reverse(name = "Pentration depth (m)", breaks = seq(0,20,5))+
  scale_color_brewer(palette = "Set1", direction = -1)+
  labs(title = "Incremental CT and temperature changes")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.title = element_blank())

p_pene_cum <- penetration_depth %>% 
  ggplot(aes(date_time_ID, penetration_depth_cum, col=var))+
  geom_hline(yintercept = 0)+
  geom_hline(data = penetration_depth_mean,
             aes(yintercept = penetration_depth_diff_mean,
                 col=var, linetype = "mean"))+
  geom_line(aes(linetype="cruise"))+
  geom_point()+
  scale_y_reverse(name = "Pentration depth (m)", breaks = seq(0,20,5))+
  scale_color_brewer(palette = "Set1", direction = -1)+
  labs(title = "Cumulative CT and temperature changes")+
  theme(axis.title.x = element_blank(),
        legend.title = element_blank())

p_pene_diff / p_pene_cum +
  plot_layout(guides = "collect")

```

