---
title: "data_base"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r packages}
library(tidyverse)
library(data.table)
library(lubridate)
```


# CTD Sensor data

## Regular profiles and transects

Reading the CTD Sensor data including recordings from auxillary pH, O~2~, Chla and pCO~2~ data. pCO~2~ data were also internally recorded on the Contros HydroC instrument and will later be replaced.

```{r read_profile_transect_CTD_data}

setwd("C:/Mueller_Jens_Data/Research/Projects/BloomSail/data/TinaV/Sensor/Profiles_Transects")
files <- list.files(pattern = "[.]cnv$")
#file <- files[1]

for (file in files){
  
  start.date <- data.table(read.delim(file, sep="#", nrows = 160))[[78,1]]
  start.date <- substr(start.date, 15, 34)
  start.date <- mdy_hms(start.date, tz="UTC")
  
  tempo <- read.delim(file, sep="", skip = 160, header = FALSE)
  tempo <- data.table(tempo[,c(2,3,4,5,6,7,9,11,13)])
  names(tempo) <- c("date", "Dep.S", "Tem.S", "Sal.S", "V_pH", "pH", "Chl", "O2", "pCO2")
  tempo$start.date <- start.date
  tempo$date <- tempo$date + tempo$start.date
  
  tempo$transect.ID <- substr(file, 1, 6)
  tempo$type <- substr(file, 8,8)
  tempo$label <- substr(file, 8,10)
  
  tempo$cast <- "up"
  tempo[date < mean(tempo[Dep.S == max(tempo$Dep.S)]$date)]$cast <- "down"
  
  if (exists("dataset")){
    dataset <- rbind(dataset, tempo)
  }
  
  if (!exists("dataset")){
    dataset <- tempo
  }
  
  rm(start.date)
  rm(tempo)
}


df <- dataset
rm(dataset, file, files)


#### plots to check succesful read-in and data-quality ####
#### removal of errornous recordings ####

#### Profiling data

# Temperature

# df %>% 
#   filter(type == "P") %>% 
#   ggplot(aes(Tem.S, Dep.S, col=label, linetype = cast))+
#   geom_line()+
#   scale_y_reverse()+
#   geom_vline(xintercept = c(10, 20))+
#   facet_wrap(~transect.ID)

df[transect.ID == "180723" & label == "P07" & Dep.S < 2 & cast == "up"]$Tem.S <- NA

# Salinity

# df %>% 
#   filter(type == "P") %>% 
#   ggplot(aes(Sal.S, Dep.S, col=label, linetype = cast))+
#   geom_path()+
#   scale_y_reverse()+
#   facet_wrap(~transect.ID)

df[Sal.S < 6]$Sal.S <- NA

# pH

# df %>% 
#   filter(type == "P") %>% 
#   ggplot(aes(pH, Dep.S, col=label, linetype=cast))+
#   geom_path()+
#   scale_y_reverse()+
#   facet_wrap(~transect.ID)
# 
# df %>% 
#   filter(type == "P") %>% 
#   ggplot(aes(V_pH, Dep.S, col=label, linetype=cast))+
#   geom_path()+
#   scale_y_reverse()+
#   facet_wrap(~transect.ID)

df[pH < 7.5]$V_pH <- NA
df[pH < 7.5]$pH <- NA

df[transect.ID == "180709" & label == "P03" & Dep.S < 5 & cast == "down"]$pH <- NA
df[transect.ID == "180709" & label == "P05" & Dep.S < 10 & cast == "down"]$pH <- NA
df[transect.ID == "180718" & label == "P10" & Dep.S < 3 & cast == "down"]$pH <- NA
df[transect.ID == "180815" & label == "P03" & Dep.S < 2 & cast == "down"]$pH <- NA
df[transect.ID == "180820" & label == "P11" & Dep.S < 15 & cast == "down"]$pH <- NA

df[transect.ID == "180709" & label == "P03" & Dep.S < 5 & cast == "down"]$V_pH <- NA
df[transect.ID == "180709" & label == "P05" & Dep.S < 10 & cast == "down"]$V_pH <- NA
df[transect.ID == "180718" & label == "P10" & Dep.S < 3 & cast == "down"]$V_pH <- NA
df[transect.ID == "180815" & label == "P03" & Dep.S < 2 & cast == "down"]$V_pH <- NA
df[transect.ID == "180820" & label == "P11" & Dep.S < 15 & cast == "down"]$V_pH <- NA


# pCO2

# df %>% 
#   filter(type == "P") %>% 
#   ggplot(aes(pCO2, Dep.S, col=label, linetype = cast))+
#   geom_path()+
#   scale_y_reverse()+
#   facet_wrap(~transect.ID)

df[transect.ID == "180616"]$pCO2 <- NA

# O2

# df %>% 
#   filter(type == "P") %>% 
#   ggplot(aes(O2, Dep.S, col=label, linetype = cast))+
#   geom_path()+
#   scale_y_reverse()+
#   facet_wrap(~transect.ID)

# Chlorophyll

# df %>% 
#   filter(type == "P") %>% 
#   ggplot(aes(Chl, Dep.S, col=label, linetype = cast))+
#   geom_path()+
#   scale_y_reverse()+
#   facet_wrap(~transect.ID)

df[Chl > 100]$Chl <- NA


#### Surface transect data

# df %>% 
#   filter(type == "T") %>% 
#   ggplot(aes(date, Dep.S, col=label))+
#   geom_point()+
#   scale_y_reverse()+
#   facet_wrap(~transect.ID, scales = "free_x")
# 
# df %>% 
#   filter(type == "T") %>% 
#   ggplot(aes(date, Tem.S, col=label))+
#   geom_point()+
#   facet_wrap(~transect.ID, scales = "free_x")
# 
# df %>% 
#   filter(type == "T") %>% 
#   ggplot(aes(date, Sal.S, col=label))+
#   geom_point()+
#   facet_wrap(~transect.ID, scales = "free_x")
# 
# df %>% 
#   filter(type == "T") %>% 
#   ggplot(aes(date, pCO2, col=label))+
#   geom_point()+
#   facet_wrap(~transect.ID, scales = "free_x")
# 
# df %>% 
#   filter(type == "T") %>% 
#   ggplot(aes(date, pH, col=label))+
#   geom_point()+
#   facet_wrap(~transect.ID, scales = "free_x")
# 
# df %>% 
#   filter(type == "T") %>% 
#   ggplot(aes(date, Chl, col=label))+
#   geom_point()+
#   facet_wrap(~transect.ID, scales = "free_x")

df[type == "T" & Chl > 10]$Chl <- NA


# df %>% 
#   filter(type == "T") %>% 
#   ggplot(aes(date, O2, col=label))+
#   geom_point()+
#   facet_wrap(~transect.ID, scales = "free_x")

df <-
  df %>% 
  select(date_time=date,
         ID=transect.ID,
         type,
         station=label,
         cast,
         dep=Dep.S,
         sal=Sal.S,
         tem=Tem.S,
         pCO2,pH,V_pH,O2,Chl)

df %>% 
  write_csv(here::here("data/_summarized_data_files", "Tina_V_Sensor_Profiles_Transects.csv"))

```

```{r Overview_temperature_profiles}
# df %>% 
#   filter(type == "P") %>% 
#   ggplot(aes(Sal.S, Dep.S, col=label, linetype = cast))+
#   geom_path()+
#   scale_y_reverse()+
#   facet_wrap(~transect.ID)

```



## Around Ostergarnsholm island

## Crossing large vessels 

# pCO~2~ data

# Merging observations

## pCO~2~ & CTD

### Time lag correction

### Temporal interpolation

