---
title: "Merging and interpolation of observations"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r packages}
library(tidyverse)
library(lubridate)
library(zoo)

```

```{r read_parameters, include = FALSE}
parameters <-
  read_rds(here::here("data",
                       "parameters.rds"))
```


```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# CTD (ts) + HydroC CO~2~ data (th, V1)

## Merging summarized data sets

Before merging the ts and th data set, the time stamp of ts is adjusted to match exactly that of th, based on zeroing pCO~2~ values recorded in the analog output (ts) and the internal memory (th).

The th data set used here, is V1 of the post-processed data by KM Contros.

```{r load_summarized_data_synchronize_time_stamp}

# Load Sensor and HydroC data ---------------------------------------------
ts <- read_csv(here::here("data/intermediate/_summarized_data_files",
                          "ts.csv"),
               col_types = list("pCO2_analog" = col_double()))

th <- read_csv(here::here("data/intermediate/_summarized_data_files",
                          "th.csv"))

# Time offset correction ----------------------------------------------

# Time offset was determined by comparing zeroing reads from Sensor and th
# in the plots produced in the section Time stamp synchronicity below
# before applying this correction

ts <- ts %>% 
  mutate(day = yday(date_time),
         date_time = if_else(day >= 206 & day <= 220,
                             date_time - 80, date_time - 10)) %>% 
  select(-day)

# Merge Sensor and HydroC data --------------------------------------------

ts_th <- full_join(ts, th) %>% 
  arrange(date_time)

rm(th, ts)
```

A pdf with plots of the zeroing signals to check the time stamp syncronicity, can be found here:

[Link to zeroing plots](https://github.com/jens-daniel-mueller/BloomSail/tree/master/output/Plots/merging_interpolation/Zero_time_synchronization.pdf){target="_blank"}


## Interpolation to common time stamp

CTD (ts) and auxillary recordings (15 sec measurment interval) are interpolated to the HydroC (th) time stamps (first 10 sec, than 1 sec measurement interval). Interpolation of ts data is not done when gaps between observations are larger than 20, indicating that th was running without ts, eg during data download from th. Thereafter, th readings not falling in regular transects/profilings are removed, by removing rows with NA depth values. Furthermore, ts readings without corresponding th readings are removed, except during periods when th was not operating.

```{r interpolation_to_common_timestamp}

# Interpolate Sensor data to HydroC time stamp

ts_th <- ts_th %>%
  mutate(
    dep_maxgap = na.approx(dep, na.rm = FALSE, maxgap = 20),
    dep = approxfun(date_time, dep)(date_time),
    sal = approxfun(date_time, sal)(date_time),
    tem = approxfun(date_time, tem)(date_time),
    pCO2_analog = approxfun(date_time, pCO2_analog)(date_time)
  ) %>%
  #remove HC readings not falling in regular transects/profiling
  filter(!is.na(dep_maxgap)) %>% 
  select(-dep_maxgap) %>%
  fill(ID, type, station) %>%
  # removes CTD readings without corresponding HydroC reading
  filter(!is.na(deployment),!is.na(pCO2_analog))


```

```{r time_stamp_synchronicity_check, eval=FALSE}
# Time stamp synchronicity

ts_th_Zero <- ts_th %>%
  filter(Zero == 1 | Flush == 1 & duration < 120)

pdf(
  file = here::here(
    "output/Plots/merging_interpolation",
    "Zero_time_synchronization.pdf"
  ),
  onefile = TRUE,
  width = 5,
  height = 5
)

for (i_deployment in unique(ts_th$deployment)) {
  #i_deployment <- unique(ts_th_Zero$deployment)[1]
  
  ts_th_Zero_deployment <- ts_th_Zero %>%
    filter(deployment == i_deployment)
  
  for (i_Zero_counter in unique(ts_th_Zero_deployment$Zero_counter)) {
    #i_Zero_counter <- unique(ts_th_Zero_deployment$Zero_counter)[1]
    
    print(
      ts_th_Zero_deployment %>%
        filter(Zero_counter == i_Zero_counter) %>%
        ggplot() +
        geom_point(aes(date_time, pCO2_corr, col = "HydroC")) +
        geom_point(aes(date_time, pCO2_analog, col = "analog")) +
        labs(
          title = paste("Depl: ", i_deployment,
                        " | Zero_counter: ", i_Zero_counter)
        )
      
    )
    
  }
}

dev.off()

rm(ts_th_Zero,
   ts_th_Zero_deployment,
   i_deployment,
   i_Zero_counter)

```

## Read cleaned, processed pCO~2~ data (V2)

A revised post-processed HydroC pCO~2~ data set was provided by KM Contros after applying a drift correction to the cleaned raw data, i.e. those without data recorded during configuration and testing of the sensor. This data set is referred to as V2. The post-processing was still based on pre- and post-deployment calibration results. 

```{r read_th_post_processed_by_manufacturer_without_AW}

# Read Contros corrected data file, based on cleaned recordings and
# without water vapor correction

th_new_withoutAW_all <-
  read_csv2(
    here::here(
      "data/input/TinaV/Sensor/HydroC-pCO2/corrected_Contros",
      "parameter&pCO2s(method 43)_new_withoutAW.txt"
    ),
    col_names = c(
      "date_time",
      "Zero",
      "Flush",
      "p_NDIR",
      "p_in",
      "T_control",
      "T_gas",
      "%rH_gas",
      "Signal_raw",
      "Signal_ref",
      "T_sensor",
      "pCO2_corr",
      "Runtime",
      "nr.ave"
    )
  ) %>%
  mutate(
    date_time = dmy_hms(date_time),
    Flush = as.factor(as.character(Flush)),
    Zero = as.factor(as.character(Zero))
  )

# slive every 10th data point to reduce number for plotting
th_new_withoutAW <- th_new_withoutAW_all %>%
  slice(seq(1, n(), 10))

# load analog pCO2 data (raw)
th_pre_cleaning <-
  read_csv(here::here(
    "data/intermediate/_summarized_data_files",
    "th_pre_cleaning.csv"
  ))

# slive every 10th data point to reduce number for plotting
th_pre_cleaning <- th_pre_cleaning %>%
  slice(seq(1, n(), 10))


# slive every 10th data point to reduce number for plotting
ts_th_sub <- ts_th %>%
  slice(seq(1, n(), 10))

```

```{r read_th_post_processed_by_manufacturer_with_AW, include=FALSE}

# Read Contros corrected data file, based on cleaned recordings and 
# with water vapor (AW) correction
# the AW correction was tested, but not used in this study

th_new_withAW <-
  read_csv2(
    here::here(
      "data/input/TinaV/Sensor/HydroC-pCO2/corrected_Contros",
      "parameter&pCO2s(method 43)_new_withAW.txt"
    ),
    col_names = c(
      "date_time",
      "Zero",
      "Flush",
      "p_NDIR",
      "p_in",
      "T_control",
      "T_gas",
      "%rH_gas",
      "Signal_raw",
      "Signal_ref",
      "T_sensor",
      "pCO2_corr",
      "Runtime",
      "nr.ave"
    )
  ) %>%
  mutate(
    date_time = dmy_hms(date_time),
    Flush = as.factor(as.character(Flush)),
    Zero = as.factor(as.character(Zero))
  )

th_new_withAW <- th_new_withAW %>%
  slice(seq(1, n(), 10))

```

### Comparison of pCO~2~ time series

#### Analog vs internal

Below, pCO~2~ time series for the analog output and the post-processed (ie drift corrected) data V1 are shown.

```{r pCO2_time_series, fig.cap="pCO~2~ record after interpolation to HydroC timestamp (analog output from HydroC recorded with CTD and drift corrected HydroC data V1 provided by Contos). ID refers to the starting date of each cruise. Please note that measurement range of the analog signal is technically restricted to 100-500  µatm. Zeroing periods are included.", fig.asp = 3}

ggplot() +
  geom_path(data = ts_th_sub,
            aes(date_time, pCO2_corr, col = "HydroC V1")) +
  geom_path(data = ts_th_sub,
            aes(date_time, pCO2_analog, col = "analog CTD")) +
  scale_color_brewer(palette = "Set1", name = "pCO2 record") +
  coord_cartesian(ylim = c(0, 600)) +
  labs(y = expression(pCO[2] ~ (µatm)), x = "") +
  facet_wrap(~ deployment, scales = "free_x", ncol = 1)

```

#### Internal V1 vs. V2

Below, pCO~2~ time series for the post-processed (ie drift corrected) data V1 and V2 are shown.

```{r compare_processed_hc_files, fig.asp=3}

th_comparison <- full_join(
  ts_th_sub %>% select(date_time, deployment, pCO2_corr),
  th_new_withAW %>% select(date_time, pCO2_corr) %>% rename(pCO2_withAW = pCO2_corr)
)

th_comparison <- full_join(
  th_comparison,
  th_new_withoutAW %>% select(date_time, pCO2_corr) %>% rename(pCO2_withoutAW = pCO2_corr)
)

th_comparison %>%
  ggplot() +
  geom_path(aes(date_time, pCO2_corr, col = "V1")) +
  geom_path(aes(date_time, pCO2_withoutAW, col = "V2")) +
  scale_color_brewer(palette = "Set1", name = "HydroC pCO2 (th)") +
  coord_cartesian(ylim = c(0, 600)) +
  labs(y = expression(pCO[2] ~ (µatm)), x = "") +
  facet_wrap( ~ deployment, scales = "free_x", ncol = 1)

```

```{r water_vapor_correction, fig.asp=5, eval=FALSE, include=FALSE}

th_comparison %>%
  ggplot() +
  geom_path(data = th_pre_cleaning %>% slice_sample(prop = 0.1),
            aes(date_time, 0, col = "pre runtime")) +
  geom_path(aes(date_time, pCO2_corr - pCO2_withAW, col = "orig - with AW")) +
  scale_color_brewer(palette = "Set1", name = "pCO2 record") +
  labs(y = expression(pCO[2] ~ (µatm)), x = "") +
  facet_wrap( ~ deployment, scales = "free_x", ncol = 1)

```

```{r water_vapor_correction_offset, fig.asp=5, eval=FALSE, include=FALSE}

th_comparison %>%
  ggplot() +
  geom_path(data = th_pre_cleaning, aes(date_time, 0, col = "pre runtime")) +
  geom_path(aes(date_time, pCO2_withoutAW - pCO2_withAW, col = "without - with AW")) +
  scale_color_brewer(palette = "Set1", name = "pCO2 record") +
  labs(y = expression(pCO[2] ~ (µatm)), x = "") +
  facet_wrap( ~ deployment, scales = "free_x", ncol = 1)

rm(ts_th_sub,
   th_pre_cleaning,
   th_new_withAW,
   th_new_withoutAW,
   th_comparison)

```


### Replace pCO~2~ data

The pCO~2~ data V1 are replaced with V2, in the data set that will be used for further analysis.

```{r replace_with_clean_pCO2_record}

th_new_withoutAW_all <- th_new_withoutAW_all %>% 
  select(date_time, pCO2_corr)

ts_th <- ts_th %>% 
  select(-pCO2_corr)

ts_th <- full_join(ts_th, th_new_withoutAW_all) 

rm(th_new_withoutAW_all)
```


### Offset analog vs post-processed pCO~2~ (V2)

Below, the offset pCO~2~ time series between the analog output and the post-processed (ie drift corrected) data V2 are shown, which allows to judge the magnitude of the drift correction applied to V2.

```{r pCO2_diff_time_series, fig.cap="pCO~2~ difference between analog and drift corrected data provided by Contos. Please note that pCO2 range is restricted to 0 - 30 µatm.", fig.asp = 3}

ts_th %>%
  ggplot() +
  geom_path(aes(date_time, pCO2_corr - pCO2_analog)) +
  ylim(-30, 0) +
  labs(y = expression(pCO[2] ~ (µats_th)), x = "") +
  facet_wrap( ~ deployment, scales = "free_x", ncol = 1)

```


# Merge sensor (ts_th) + track (tt) data

```{r merge_sensor_track}

tt <- read_csv(here::here("data/intermediate/_summarized_data_files",
                          "tt.csv"))


tm <- full_join(ts_th, tt) %>% 
  arrange(date_time)

# interpolate tt data and than remove columns that originate from tt time stamp
tm <- tm %>%
  mutate(lat = approxfun(date_time, lat)(date_time),
         lon = approxfun(date_time, lon)(date_time)) %>% 
  filter(!is.na(dep))


```


# Write merged file

```{r write_merged_data_file}


tm %>% write_csv(here::here("data/intermediate/_merged_data_files/merging_interpolation",
                            "tm.csv"))

rm(tm, ts_th, tt)

```

