---
title: "Merging and interpolation of observations"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```


```{r packages}
library(tidyverse)
library(lubridate)
library(zoo)

```



```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# CTD (ts) + HydroC CO~2~ data (th)

## Merging summarized data sets


```{r load_summarized_data_synchronize_time_stamp, eval=FALSE}

# Load Sensor and HydroC data ---------------------------------------------
ts <- read_csv(here::here("Data/_summarized_data_files",
                          "ts.csv"),
               col_types = list("pCO2_analog" = col_double()))

th <- read_csv(here::here("Data/_summarized_data_files",
                          "th.csv"))

# Time offset correction ----------------------------------------------

# Time offset was determined by comparing zeroing reads from Sensor and th
# in the plots produced in the section Time stamp synchronzity below
# before applying this correction

ts <- ts %>% 
  mutate(day = yday(date_time),
         date_time = if_else(day >= 206 & day <= 220,
                             date_time - 80, date_time - 10)) %>% 
  select(-day)

# Merge Sensor and HydroC data --------------------------------------------

ts_th <- full_join(ts, th) %>% 
  arrange(date_time)

# ts_th_full <- full_join(ts, th_full) %>% 
#   arrange(date_time)

rm(th, ts)
```


## Interpolation to common time stamp

CTD and auxillary recordings (15 sec measurment interval) are interpolated to HydroC time stamps (first 10 sec, than 1 sec measurement interval) when gaps between observations are not larger than 20. Thereafter, HydroC readings not falling in regular transects/profilings are removed, by removing rows with NA depth values. Furthermore, CTD readings without corresponding HydroC reading are removed, except during periods when HydroC was not operating.

```{r interpolation_to_common_timestamp, eval=FALSE}

# Interpolate Sensor data to HydroC timestamp

ts_th <- ts_th %>%
  mutate(dep_maxgap = na.approx(dep, na.rm = FALSE, maxgap = 20),
         dep = approxfun(date_time, dep)(date_time),
         sal = approxfun(date_time, sal)(date_time),
         tem = approxfun(date_time, tem)(date_time),
         pCO2_analog = approxfun(date_time, pCO2_analog)(date_time)) %>% 
  filter(!is.na(dep_maxgap)) %>% #remove HC readings not falling in regular transects/profilings
  select(- dep_maxgap) %>% 
  fill(ID, type, station) %>% 
  filter(!is.na(deployment), !is.na(pCO2_analog)) # removes CTD readings without corresponding HydroC reading

  # filter(!is.na(deployment) | is.na(pCO2_analog)) # removes CTD readings without corresponding HydroC reading, except during periods when HydroC was not operating


# Time stamp synchronzity

ts_th_Zero <- ts_th %>%
  filter(Zero == 1 | Flush == 1 & duration < 120)

pdf(file=here::here("output/Plots/merging_interpolation",
                    "Zero_time_synchronization.pdf"),
    onefile = TRUE, width = 5, height = 5)

for (i_deployment in unique(ts_th$deployment)) {

  #i_deployment <- unique(ts_th_Zero$deployment)[1]

  ts_th_Zero_deployment <- ts_th_Zero %>%
    filter(deployment == i_deployment)

  for (i_Zero_counter in unique(ts_th_Zero_deployment$Zero_counter)) {

    #i_Zero_counter <- unique(ts_th_Zero_deployment$Zero_counter)[1]

    print(

    ts_th_Zero_deployment %>%
      filter(Zero_counter == i_Zero_counter) %>%
      ggplot()+
      geom_point(aes(date_time, pCO2_corr, col="HydroC"))+
      geom_point(aes(date_time, pCO2_analog, col="analog"))+
      labs(title = paste("Depl: ",i_deployment,
                         " | Zero_counter: ", i_Zero_counter))

    )

  }
}

dev.off()

rm(ts_th_Zero, ts_th_Zero_deployment, i_deployment, i_Zero_counter)

```

## Write merged file

```{r safe_merged_data_file, eval=FALSE}

ts_th %>% 
  write_csv(here::here("Data/_merged_data_files/merging_interpolation", "ts_th.csv"))
#write_csv(ts_th_full, here::here("Data/_merged_data_files", "BloomSail_CTD_HydroC_full.csv"))

rm(ts_th)

```

## Time series pCO~2~

### Absolute values

```{r pCO2_time_series, fig.cap="pCO~2~ record after interpolation to HydroC timestamp (analog output from HydroC and drift corrected data provided by Contos). ID refers to the starting date of each cruise. Please note that pCO2_analog measurement range is technically restricted to 100-500  µats_th. Zeroing periods are included.", fig.asp = 3}

ts_th <- read_csv(here::here("data/_merged_data_files/merging_interpolation",
                          "ts_th.csv"),
               col_types = cols(ID = col_character(),
                                pCO2_analog = col_double(),
                                pCO2_corr = col_double(),
                                Zero = col_factor(),
                                Flush = col_factor(),
                                Zero_counter = col_integer(),
                                deployment = col_integer(),
                                duration = col_double(),
                                mixing = col_character()))


ts_th %>% 
  filter(!is.na(pCO2_corr)) %>% 
  ggplot()+
  geom_path(aes(date_time, pCO2_corr, col = "HydroC, drift corrected"))+
  geom_path(aes(date_time, pCO2_analog, col = "analog CTD"))+
  scale_color_brewer(palette = "Set1", name = "pCO2 record")+
  coord_cartesian(ylim = c(0,600))+
  labs(y=expression(pCO[2]~(µats_th)), x="")+
  facet_wrap(~ID, scales = "free_x", ncol = 1)

```

### Offset analog vs post-processed pCO~2~

```{r pCO2_diff_time_series, fig.cap="pCO~2~ difference betweeb HydroC and drift corrected data provided by Contos. Please note that pCO2 range is restricted to +/- 50  µatm.", fig.asp = 3}

ts_th %>% 
  filter(!is.na(pCO2_corr)) %>% 
  ggplot()+
  geom_path(aes(date_time, pCO2_corr - pCO2_analog))+
  ylim(-30, 0)+
  labs(y=expression(pCO[2]~(µats_th)), x="")+
  facet_wrap(~ID, scales = "free_x", ncol = 1)

```

```{r measurement_frequency, eval=FALSE, include=FALSE}

ts_th <- ts_th %>% 
  mutate(dt = c(0,diff(date_time)))

ts_th %>% 
  filter(dt < 30) %>% 
  ggplot(aes(date_time, dt))+
  geom_point()

```

# Merges sensor (ts_th) + track (tt) data

```{r merge_sensor_track, eval=FALSE}

ts_th <- read_csv(here::here("data/_merged_data_files/merging_interpolation",
                          "ts_th.csv"),
               col_types = cols(ID = col_character(),
                                pCO2_analog = col_double(),
                                pCO2_corr = col_double(),
                                Zero = col_factor(),
                                Flush = col_factor(),
                                Zero_counter = col_integer(),
                                deployment = col_integer(),
                                duration = col_double(),
                                mixing = col_character()))


tt <- read_csv(here::here("Data/_summarized_data_files",
                          "tt.csv"))


tm <- full_join(ts_th, tt) %>% 
  arrange(date_time)

# interpolate tt data and than remove columns that originate from tt time stamp
tm <- tm %>%
  mutate(lat = approxfun(date_time, lat)(date_time),
         lon = approxfun(date_time, lon)(date_time)) %>% 
  filter(!is.na(dep))

tm %>% write_csv(here::here("Data/_merged_data_files/merging_interpolation",
                            "tm.csv"))

rm(tm, ts_th, tt)
```

# Tasks / open questions

