---
title: "NCP reconstruction"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---


```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r packages}
library(tidyverse)
library(ncdf4)
library(seacarb)
library(oce)
library(patchwork)
library(lubridate)
library(metR)
# library(marelac)
```


```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# Approach

In order to test how (and how well) the depth-integrated C~T~ estimates can be reproduced if only surface CO~2~ data were available, the BloomSail observations were restricted to those made in surface water and following reconstruction approaches were tested:

- Mixed layer depth: Integration of surface observation across the MLD, assuming homogenious vertical patterns
- C~T~ profile reconstruction: Vertical reconstruction of incremental C~T~ changes based on profiles of incremental changes in temperature
- Temperature penetration depth:  Integration of surface observation across the temperature penetration depth, assuming similar vertical extension as for C~T~ drawdown.

Note: The reconstruction of C~T~ profiles and the integration across the temperature penetration depth should produce very similar results. However, the latter avoids to create misinterpretable information about the vertical distribution of C~T~.


```{r subsetting criteria}

max_dep <- 25
surface_dep <- 6

integration_dep <- 12

```


# BloomSail data

1m gridded, downcast profiles were used.
Mean CO~2~ data from upper 6 metres were used as surface values.


```{r read_tm}

tm_profiles_ID <-
  read_csv(here::here("Data/_merged_data_files/CT_dynamics", "tm_profiles_ID.csv"))

tm_profiles_ID <- tm_profiles_ID %>% 
  select(- c(date_ID))

```

```{r calculate_cumulative_changes_tm}

tm_profiles_ID_long <- tm_profiles_ID %>%
  select(-c(pCO2, sal)) %>% 
  pivot_longer(c("tem", "nCT"), values_to = "value", names_to = "var") %>% 
  group_by(var, dep) %>%
  arrange(date_time_ID) %>%
  mutate(#date_time_ID_diff = as.numeric(date_time_ID - lag(date_time_ID)),
         #date_time_ID_ref  = date_time_ID - (date_time_ID - lag(date_time_ID))/2,
         value_diff = value     - lag(value, default = first(value)),
         #value_diff_daily = value_diff / date_time_ID_diff,
         value_cum = cumsum(value_diff)) %>% 
  ungroup()


```

## C~T~ and tem penetration depth

### Cumulative on July 23

```{r penetration_depth_cumulative_tm}

tm_profiles_ID_long_180723 <- tm_profiles_ID_long %>% 
  filter(ID == 180723)

tm_profiles_ID_long_180723_dep <- tm_profiles_ID_long_180723 %>% 
  select(var, dep, value_cum) %>% 
  mutate(value_cum = if_else(value_cum > 0 & var == "nCT",
                             NaN, value_cum),
         value_cum = if_else(value_cum < 0 & var == "tem",
                             NaN, value_cum)) %>% 
  group_by(var) %>% 
  arrange(dep) %>% 
  mutate(value_cum_i = sum(value_cum, na.rm = TRUE),
         value_cum_dep = cumsum(value_cum),
         value_cum_i_rel = value_cum_dep/value_cum_i*100) %>% 
  ungroup()

value_cum <- tm_profiles_ID_long_180723_dep %>% 
  group_by(var) %>% 
  summarise(value_cum_i = mean(value_cum_i)) %>% 
  ungroup()

value_surface <- tm_profiles_ID_long_180723 %>% 
  select(var, dep, value_cum) %>% 
  filter(dep < surface_dep) %>%
  group_by(var) %>% 
  summarise(value_surface = mean(value_cum)) %>% 
  ungroup()

d_pen <- full_join(value_cum, value_surface)
d_pen <- d_pen %>% 
  mutate(d_pen = value_cum_i / value_surface)
  
rm(value_cum, value_surface)


p_tm_profiles_ID_long <- tm_profiles_ID_long_180723 %>% 
  arrange(dep) %>% 
  ggplot(aes(value_cum, dep))+
  geom_hline(aes(yintercept = integration_dep, col="integration"))+
  geom_hline(data = d_pen, aes(yintercept = d_pen, col="penetration"))+
  geom_vline(xintercept = 0)+
  geom_point()+
  geom_path()+
  scale_y_reverse()+
  scale_color_brewer(palette = "Dark2", guide = FALSE)+
  labs(y = "Depth (m)", x="Cumulative change")+
  theme(legend.position = "left")+
  facet_wrap(var~., ncol = 1, scales = "free_x")

p_tm_profiles_ID_long_rel <- tm_profiles_ID_long_180723_dep %>% 
  ggplot(aes(value_cum_i_rel, dep))+
  geom_hline(aes(yintercept = integration_dep, col="integration"))+
  geom_hline(data = d_pen, aes(yintercept = d_pen, col="penetration"))+
  geom_vline(xintercept = 90)+
  geom_point()+
  geom_line()+
  scale_y_reverse(limits = c(25,0))+
  scale_color_brewer(palette = "Dark2", name = "Depth")+
  scale_x_continuous(limits = c(0, NA))+
  labs(x = "Relative contribution (%)")+
  facet_wrap(var~., ncol = 1, scales = "free_x")+
  theme(axis.title.y = element_blank())

p_tm_profiles_ID_long + p_tm_profiles_ID_long_rel

rm(tm_profiles_ID_long_180723,
   tm_profiles_ID_long_180723_dep,
   p_tm_profiles_ID_long,
   p_tm_profiles_ID_long_rel)

```

### Daily

```{r penetration_depth_daily_tm}

# surface values
diff_surface <- tm_profiles_ID_long %>% 
  filter(dep < surface_dep) %>% 
  group_by(ID, var) %>% 
  summarise(value_diff_surface = mean(value_diff, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(value_diff_surface = if_else(value_diff_surface > 0 & var == "nCT",
                                      NaN, value_diff_surface),
         value_diff_surface = if_else(value_diff_surface < 0 & var == "tem",
                                      NaN, value_diff_surface))

tm_profiles_ID_long <- full_join(tm_profiles_ID_long, diff_surface) 
rm(diff_surface)

# calculate penetration depths

penetration_depth <- tm_profiles_ID_long %>% 
  mutate(value_diff = if_else(value_diff > 0 & var == "nCT",
                                      NaN, value_diff),
         value_diff = if_else(value_diff < 0 & var == "tem",
                                      NaN, value_diff)) %>% 
  group_by(var, ID, date_time_ID) %>% 
  summarise(value_diff_int = sum(value_diff, na.rm = TRUE),
            value_diff_surface = mean(value_diff_surface, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(penetration_depth_diff = value_diff_int / value_diff_surface)

penetration_depth_mean <- penetration_depth %>% 
  group_by(var) %>% 
  summarise(penetration_depth_diff_mean = mean(penetration_depth_diff, na.rm=TRUE),
            penetration_depth_diff_sd = sd(penetration_depth_diff, na.rm=TRUE)) %>% 
  ungroup()

penetration_depth_mean

p_surface <- penetration_depth %>% 
  ggplot(aes(date_time_ID, value_diff_surface))+
  geom_hline(yintercept = 0)+
  geom_line()+
  geom_point()+
  scale_y_reverse(name = "Change surface value")+
  scale_x_datetime(breaks = "week", date_labels = "%d %b")+
  scale_color_brewer(palette = "Set1", direction = -1)+
  theme(axis.title.x = element_blank(),
        legend.title = element_blank())+
  facet_grid(var~., scales = "free_y")

p_integrated <- penetration_depth %>% 
  ggplot(aes(date_time_ID, value_diff_int))+
  geom_hline(yintercept = 0)+
  geom_line()+
  geom_point()+
  scale_y_reverse(name = "Change integrated value")+
  scale_x_datetime(breaks = "week", date_labels = "%d %b")+
  scale_color_brewer(palette = "Set1", direction = -1)+
  theme(axis.title.x = element_blank(),
        legend.title = element_blank())+
  facet_grid(var~., scales = "free_y")

p_pen_dep <- penetration_depth %>% 
  ggplot(aes(date_time_ID, penetration_depth_diff, col=var))+
  geom_hline(yintercept = 0)+
  geom_hline(data = penetration_depth_mean,
             aes(yintercept = penetration_depth_diff_mean,
                 col=var, linetype = "mean"))+
  geom_line(aes(linetype="cruise"))+
  geom_point()+
  scale_y_reverse(name = "Penetration depth (m)", breaks = seq(0,20,5))+
  scale_x_datetime(breaks = "week", date_labels = "%d %b")+
  scale_color_brewer(palette = "Set1", direction = -1)+
  theme(axis.title.x = element_blank(),
        legend.title = element_blank())


p_surface + p_integrated + p_pen_dep + 
  plot_layout(ncol = 1)

rm(p_surface, p_integrated, p_pen_dep)
rm(penetration_depth, penetration_depth_mean, tm_profiles_ID_long)

```


## Finnmaid

```{r gt_subsetting_criteria}

# route
select_route <- "E"

# latitude limits
low_lat <- 57.3
high_lat <- 57.5

# date limits
start_date <- "2018-06-20"
end_date <- "2018-08-25"

fixed_values <- 
  read_csv(here::here("Data/_summarized_data_files", "tb_fix.csv"))


```

### Approach

In order to test how (and how well) the depth-integrated C~T~ estimates can be reproduced if only surface CO~2~ data from VOS Finnamaid and hydrographical GETM model date were available, two reconstruction approaches were tested:

1. MLD: Integration of surface observation across the MLD, assuming homogenious vertical patterns
2. Temperature: Vertical reconstruction of incremental CT changes based on profiles of incremental changes in temperature

### GETM

The following information refer to regional mean values modeled along the Finnmaid track east and within the BloomSail field work area (57.3 - 57.5 N). 

## STD Profiles 

Mean daily salinity and temperature profiles were extracted from 3d GETM data (vertically resolved transects).

```{r read_gt_3d_sal_tem_age, eval=FALSE}

nc <- nc_open(here::here("data/GETM", "Finnmaid.E.3d.2018.nc"))
#print(nc$var)

lat <- ncvar_get(nc, "latc")

time_units <- nc$dim$time$units %>%     #we read the time unit from the netcdf file to calibrate the time 
    substr(start = 15, stop = 33) %>%   #calculation, we take the relevant information from the string
    ymd_hms()                           # and transform it to the right format
t <- time_units + ncvar_get(nc, "time") # read time vector
rm(time_units)

d <- ncvar_get(nc, "zax") # read depths vector

for (var_3d in c("salt", "temp", "SurfaceAge")) {
  
array <- ncvar_get(nc, var_3d) # store the data in a 3-dimensional array
#dim(array) # should be 3d with dimensions: 544 coordinates, 51 depths, and number of days of month

fillvalue <- ncatt_get(nc, var_3d, "_FillValue")

# Working with the data
array[array == fillvalue$value] <- NA

    for (i in seq(1,length(t),1)){
      
      # i <- 3
      array_slice <- array[, , i] # slices data from one day
      
      array_slice_df <- as.data.frame(t(array_slice))
      array_slice_df <- as_tibble(array_slice_df)
      
      gt_3d_part <- array_slice_df %>%
        set_names(as.character(lat)) %>%
        mutate(dep = -d) %>%
        gather("lat", "value", 1:length(lat)) %>%
        mutate(lat = as.numeric(lat)) %>%
        filter(lat > low_lat, lat < high_lat,
               dep <= max_dep) %>%
        mutate(var = var_3d,
               date_time=t[i]) %>% 
        select(date_time, dep, value, var)

      
      if (exists("gt_3d")) {
        gt_3d <- bind_rows(gt_3d, gt_3d_part)
        } else {gt_3d <- gt_3d_part}
      
  rm(array_slice, array_slice_df, gt_3d_part)
      
    }
rm(array, fillvalue)

}

nc_close(nc)
rm(nc)

gt_3d_long <- gt_3d %>% 
  filter(date_time >= start_date & date_time <= end_date) %>% 
  group_by(date_time, var, dep) %>% 
  summarise_all(list(value=~mean(.,na.rm=TRUE))) %>% # regional averaging
  ungroup()

gt_3d_long %>% 
  write_csv(here::here("data/_summarized_data_files", file = "gt_3d_long.csv"))


rm(gt_3d, gt_3d_long, i, lat, d, t, var_3d)

```

Seawater density was calculated according to TEOS-10.

```{r calculate_rho_gt_3d}

gt_3d_long <- 
  read_csv(here::here("data/_summarized_data_files", "gt_3d_long.csv"))

gt_3d <- gt_3d_long %>% 
  pivot_wider(values_from = value, names_from = var) %>% 
  select(-SurfaceAge) %>% 
  rename(sal = salt, tem=temp)

```

```{r gt_profiles_sal_tem}

gt_3d_long <- gt_3d %>% 
  pivot_longer(c(sal, tem), values_to = "value", names_to = "var")

gt_3d_long %>% 
  ggplot(aes(value, dep,
             col=date_time,
             group=date_time))+
  geom_path()+
  scale_y_reverse(expand = c(0,0))+
  scale_color_viridis_c(name="Date", trans = "time")+
  facet_wrap(~var, scales = "free_x", ncol = 2)

rm(gt_3d_long)

```




# Finnmaid

## Data preparation

Finnmaid data, including reconstructed data during LICOS operation failure.

```{r read_Finnmaid}

fm <-
 read_csv(here::here("Data/_summarized_data_files",
                      "fm_bloomsail.csv"))

fm <- fm %>% 
  filter(date_time > start_date,
         date_time < end_date) %>% 
  select(ID, date_time, sensor, sal, tem, pCO2) %>% 
  mutate(ID = as.factor(ID))

```

### C~T~ calculation

Calculate nC~T~ based on fixed A~T~ and salinity mean values. 

```{r CT_calculation}

fm <- fm %>% 
  mutate(nCT = carb(24,
                   var1=pCO2,
                   var2=fixed_values$AT*1e-6,
                   S=fixed_values$sal,
                   T=tem,
                   k1k2="m10", kf="dg", ks="d", gas="insitu")[,16]*1e6)

```

### Regional averaging

Calculate regional mean and sd values for each crossing of the area.

```{r timeseries_of_regional_averages}

fm_ID <- fm %>% 
  pivot_longer(c(pCO2, sal, tem, nCT), values_to = "value", names_to = "var") %>% 
  group_by(ID) %>% 
  mutate(date_time_ID = mean(date_time)) %>%
  ungroup() %>% 
  select(-date_time) %>% 
  group_by(ID, date_time_ID, sensor, var) %>% 
  summarise_all(list(~mean(.), ~sd(.)), na.rm=TRUE) %>%
  ungroup() %>% 
  rename(value=mean)

```

### Read tm profile data

Read original profile data and calculate surface mean and sd values.

```{r tm_long_surface}

tm_profiles <-
  read_csv(here::here("data/_merged_data_files/CT_dynamics", "tm_profiles.csv"))

tm_profiles_ID_long_surface <- tm_profiles %>% 
  filter(dep < surface_dep) %>% 
  select(-c(dep, date_ID, station, date_time, lat, lon, pCO2_corr)) %>% 
  mutate(ID = as.factor(ID)) %>% 
  pivot_longer(sal:nCT, values_to = "value", names_to = "var") %>% 
  group_by(ID, date_time_ID, var) %>% 
  summarise_all(list(~mean(.), ~sd(.)), na.rm = TRUE) %>% 
  ungroup()


```

### Timeseries

```{r fm_tm_time_series, fig.asp=1.2}

fm_ID %>% 
  ggplot()+
  geom_rect(data = fixed_values, aes(xmin=start, xmax=end, ymin=-Inf, ymax=Inf), alpha=0.2)+
  geom_path(aes(x=date_time_ID, y=value))+
  geom_ribbon(aes(x=date_time_ID, y=value, ymax=value+sd, ymin=value-sd, fill="Finnmaid"), alpha=0.3)+
  geom_ribbon(data = tm_profiles_ID_long_surface,
             aes(x=date_time_ID, ymin=mean-sd, ymax=mean+sd, fill="BloomSail"), alpha=0.3)+
  geom_point(aes(x=date_time_ID, y=value, col=sensor))+
  geom_point(data = tm_profiles_ID_long_surface,
             aes(x=date_time_ID, y=mean, col="BloomSail"))+
  geom_line(data = tm_profiles_ID_long_surface,
             aes(x=date_time_ID, y=mean, col="BloomSail"))+
  facet_grid(var~., scales = "free_y")+
  scale_color_brewer(palette = "Set1")+
  scale_fill_brewer(palette = "Set1", name="+/- SD")+
  scale_x_datetime(date_breaks = "week",
                   date_labels = "%b %d")+
  theme(axis.title.x = element_blank())

```

## Missing observations

The observational gaps in the Finnmaid SST and C~T~ time series were filled with:

- two BloomSail observations
- an interpolated finnmaid value to match the starting date

The time series was restricted to the period where BloomSail observations are available.

```{r fm_interpolate_starting_values}

tm_start_date <- tm_profiles_ID_long_surface %>% 
  filter(ID %in% c("180705"),
         var %in% c("tem", "nCT")) %>% 
  select(date_time_ID, ID, var) %>% 
  mutate(sensor = "interpolated")

fm_tm_ID <- full_join(fm_ID, tm_start_date) %>% 
  arrange(date_time_ID) %>% 
  filter(var %in% c("tem", "nCT"))

fm_tm_ID <- fm_tm_ID %>% 
  group_by(var) %>% 
  mutate(value = approxfun(date_time_ID, value)(date_time_ID)) %>% 
  ungroup()

rm(tm_start_date)
```


```{r fm_fill_gap_with_tm}

tm_gap <- tm_profiles_ID_long_surface %>% 
  filter(ID %in% c("180718", "180723"),
         var %in% c("tem", "nCT")) %>% 
  select(date_time_ID, ID, var, value = mean) %>% 
  mutate(sensor = "BloomSail")

fm_tm_ID <- full_join(fm_tm_ID, tm_gap) %>% 
  arrange(date_time_ID) %>% 
  select(-sd) %>% 
  filter(var %in% c("tem", "nCT")) %>% 
  mutate(period = "BloomSail",
         period = if_else(date_time_ID < fixed_values$start, "pre-BloomSail", period),
         period = if_else(date_time_ID > fixed_values$end, "post-BloomSail", period))


fm_tm_ID <- fm_tm_ID %>% 
  filter(period == "BloomSail") %>% 
  select(-period)

rm(fm_ID, fm, tm_gap, tm_profiles_ID_long_surface, tm_profiles)

```

```{r fm_time_series_filled}

fm_tm_ID %>% 
  ggplot()+
  geom_path(aes(date_time_ID, value))+
  geom_point(aes(date_time_ID, value, col=sensor))+
  facet_grid(var~., scales = "free_y")+
  scale_color_brewer(palette = "Set1")+
  scale_x_datetime(date_breaks = "week",
                   date_labels = "%b %d")+
  theme(axis.title.x = element_blank())

```


# Merge all data sets

## Interpolate gt dep grid

```{r interpolate_gt_3d_dep_grid}

gt_3d_int <- gt_3d %>% 
  mutate(dep_int = dep + 0.5) %>% 
  group_by(date_time) %>% 
  mutate(sal_int = approxfun(dep, sal)(dep_int),
         tem_int = approxfun(dep, tem)(dep_int)) %>% 
  ungroup() %>% 
  select(date_time, dep=dep_int, sal=sal_int, tem=tem_int) %>% 
  drop_na()

rm(gt_3d)

```

## Merge fm and gt

```{r fm_gt_merge}

fm_tm_ID_wide <- fm_tm_ID %>% 
  filter(var %in% c("nCT")) %>% 
  select(date_time_ID, var, value) %>% 
  pivot_wider(values_from = value, names_from = var)



fm_gt <- expand_grid(fm_tm_ID_wide, dep = unique(gt_3d_int$dep))

fm_gt <- full_join(fm_gt,
                   gt_3d_int %>% rename(date_time_ID = date_time)) %>% 
  arrange(date_time_ID)

rm(fm_tm_ID_wide, gt_3d_int)

```

## Interpolate gt time stamp

```{r fm_gt_interpolate_time_stamp}

fm_gt <- fm_gt %>% 
  arrange(date_time_ID) %>% 
  group_by(dep) %>% 
  mutate(tem = approxfun(date_time_ID, tem)(date_time_ID),
         sal = approxfun(date_time_ID, sal)(date_time_ID)) %>% 
  ungroup() %>% 
  arrange(dep) %>% 
  filter(!is.na(nCT))


```

## Bind tm and fm_gt

```{r tm_fm_gt_bind}

tm_profiles_ID <- tm_profiles_ID %>% 
  select(-c(ID, pCO2)) %>% 
  mutate(source = "tm")

fm_gt <- fm_gt %>% 
  mutate(source = "fm")

tm_fm_gt <- bind_rows(tm_profiles_ID, fm_gt)

```

```{r tm_fm_gt_time_series}

tm_fm_gt_long <- tm_fm_gt %>% 
  pivot_longer(sal:nCT, values_to = "value", names_to = "var")

tm_fm_gt_long %>%
  filter(dep == 3.5) %>% 
  ggplot(aes(date_time_ID, value, col=source))+
  geom_path()+
  geom_point()+
  scale_x_datetime(date_breaks = "week",
                   date_labels = "%b %d")+
  facet_grid(var~., scales = "free_y")+
  theme(axis.title.x = element_blank())

```

```{r tm_fm_gt_hovmoeller_tem}

bin <- 2

tm_fm_gt %>% 
  ggplot(aes(date_time_ID, dep, z=tem))+
  geom_contour_fill(breaks = MakeBreaks(bin))+
  geom_vline(aes(xintercept = date_time_ID),
             col="white", linetype = "1f")+
  scale_fill_viridis_c(name="Temperature (°C)", option = "B",
                       guide = "colorstrip", breaks = MakeBreaks(bin))+
  scale_y_reverse()+
  scale_x_datetime(date_breaks = "week",
                   date_labels = "%b %d")+
  coord_cartesian(expand = 0)+
  labs(y="Depth (m)")+
  theme(axis.title.x = element_blank())+
  facet_grid(source~.)


rm(bin)
```

# Integration depths

## MLD

### Density

```{r calculate_density}

tm_fm_gt <- tm_fm_gt %>% 
  mutate(rho = swSigma(salinity = sal, temperature = tem, pressure = dep/10))

```


```{r tm_fm_gt_hovmoeller_rho}

bin <- 0.5

tm_fm_gt %>% 
  ggplot()+
  geom_contour_fill(aes(date_time_ID, dep, z=rho),
                    breaks = MakeBreaks(bin))+
  geom_vline(aes(xintercept = date_time_ID),
             col="white", linetype = "1f")+
  scale_fill_viridis_c(name="Rho", option = "B",
                       guide = "colorstrip", breaks = MakeBreaks(bin))+
  scale_y_reverse()+
  coord_cartesian(expand = 0)+
  labs(y="Depth (m)")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())+
  facet_wrap(~source)

rm(bin)
```

### MLD

```{r calculate_MLD}

tm_fm_gt_MLD <- expand_grid(tm_fm_gt, rho_lim = seq(0.1,0.5,0.1))

tm_fm_gt_MLD <- tm_fm_gt_MLD %>% 
  arrange(dep) %>% 
  group_by(date_time_ID, source, rho_lim) %>% 
  mutate(d_rho = rho - first(rho)) %>% 
  filter(d_rho > rho_lim) %>% 
  summarise(MLD = min(dep)) %>% 
  ungroup() %>% 
  mutate(rho_lim = as.factor(rho_lim))

```

```{r MLD_time_series}

tm_fm_gt_MLD %>% 
  ggplot()+
  geom_hline(yintercept = 0)+
  geom_line(aes(date_time_ID, MLD, col=rho_lim))+
  scale_y_reverse(limits = c(25,0))+
  scale_color_viridis_d(name="Rho lim")+
  labs(x="", y="MLD (m)")+
  facet_wrap(~source)


```


```{r tm_fm_gt_hovmoeller_tem_MLD}

bin <- 2

tm_fm_gt %>% 
  ggplot()+
  geom_contour_fill(aes(date_time_ID, dep, z=tem),
                    breaks = MakeBreaks(bin))+
  geom_path(data = tm_fm_gt_MLD, aes(date_time_ID, MLD, col=rho_lim))+
  scale_fill_gradient(name="Temperature (°C)",
                      guide = "colorstrip", breaks = MakeBreaks(bin),
                      high = "white",
                      low = "darkgrey")+
  scale_color_viridis_d()+
  scale_y_reverse()+
  coord_cartesian(expand = 0)+
  labs(y="Depth (m)")+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())+
  facet_wrap(~source)


rm(bin)
```


```{r filter_rho_lim}

rho_lim_value <- 0.1

tm_fm_gt_MLD <- tm_fm_gt_MLD %>% 
  filter(rho_lim == rho_lim_value) %>% 
  select(-rho_lim)


tm_fm_gt <- full_join(tm_fm_gt, tm_fm_gt_MLD)

```


## Penetration depth



```{r calculate_cumulative_changes}

tm_fm_gt_long <- tm_fm_gt %>%
  select(-c(sal)) %>% 
  pivot_longer(c("tem", "nCT"), values_to = "value", names_to = "var") %>% 
  group_by(source, var, dep) %>%
  arrange(date_time_ID) %>%
  mutate(#date_time_ID_diff = as.numeric(date_time_ID - lag(date_time_ID)),
         #date_time_ID_ref  = date_time_ID - (date_time_ID - lag(date_time_ID))/2,
         value_diff = value     - lag(value, default = first(value)),
         #value_diff_daily = value_diff / date_time_ID_diff,
         value_cum = cumsum(value_diff)) %>% 
  ungroup()

tm_fm_gt_long <- tm_fm_gt_long %>% 
  filter(var == "tem") %>% 
  select(-var)


```

## Penetration depth

### Cumulative

```{r penetration_depth_cumulative}

tm_fm_gt_long_180723 <- tm_fm_gt_long %>% 
  filter(date_time_ID == ymd_hms("2018-07-24 07:58:29"))

tm_fm_gt_long_180723_dep <- tm_fm_gt_long_180723 %>% 
  select(source, dep, value_cum) %>% 
  mutate(value_cum = if_else(value_cum < 0,
                             NaN, value_cum)) %>% 
  group_by(source) %>% 
  arrange(dep) %>% 
  mutate(value_cum_i = sum(value_cum, na.rm = TRUE),
         value_cum_dep = cumsum(value_cum),
         value_cum_i_rel = value_cum_dep/value_cum_i*100) %>% 
  ungroup()

value_cum <- tm_fm_gt_long_180723_dep %>% 
  group_by(source) %>% 
  summarise(value_cum_i = mean(value_cum_i)) %>% 
  ungroup()

value_surface <- tm_fm_gt_long_180723 %>% 
  select(source, dep, value_cum) %>% 
  filter(dep < surface_dep) %>%
  group_by(source) %>% 
  summarise(value_surface = mean(value_cum)) %>% 
  ungroup()

d_pen <- full_join(value_cum, value_surface)
d_pen <- d_pen %>% 
  mutate(d_pen = value_cum_i / value_surface)
  
rm(value_cum, value_surface)


p_tm_fm_gt_long <- tm_fm_gt_long_180723 %>% 
  arrange(dep) %>% 
  ggplot(aes(value_cum, dep))+
  geom_hline(aes(yintercept = integration_dep, col="integration"))+
  geom_hline(data = d_pen, aes(yintercept = d_pen, col="penetration"))+
  geom_vline(xintercept = 0)+
  geom_point()+
  geom_path()+
  scale_y_reverse()+
  scale_color_brewer(palette = "Dark2", guide = FALSE)+
  labs(y = "Depth (m)", x="Cumulative change")+
  theme(legend.position = "left")+
  facet_grid(.~source, scales = "free_x")

p_tm_fm_gt_long_rel <- tm_fm_gt_long_180723_dep %>% 
  ggplot(aes(value_cum_i_rel, dep))+
  geom_hline(aes(yintercept = integration_dep, col="integration"))+
  geom_hline(data = d_pen, aes(yintercept = d_pen, col="penetration"))+
  geom_vline(xintercept = 90)+
  geom_point()+
  geom_line()+
  scale_y_reverse(limits = c(25,0))+
  scale_color_brewer(palette = "Dark2", name = "Depth")+
  scale_x_continuous(limits = c(0, NA))+
  labs(x = "Relative contribution (%)")+
  facet_wrap(.~source, ncol = 1, scales = "free_x")+
  theme(axis.title.y = element_blank())

p_tm_fm_gt_long + p_tm_fm_gt_long_rel

rm(tm_fm_gt_long_180723,
   tm_fm_gt_long_180723_dep,
   p_tm_fm_gt_long,
   p_tm_fm_gt_long_rel)

```

### Daily

```{r penetration_depth_daily}

# surface values
diff_surface <- tm_fm_gt_long %>% 
  filter(dep < surface_dep) %>% 
  group_by(date_time_ID, source) %>% 
  summarise(value_diff_surface = mean(value_diff, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(value_diff_surface = if_else(value_diff_surface < 0,
                                      NaN, value_diff_surface))

tm_fm_gt_long <- full_join(tm_fm_gt_long, diff_surface) 
rm(diff_surface)

# calculate penetration depths

penetration_depth <- tm_fm_gt_long %>% 
  mutate(value_diff = if_else(value_diff < 0,
                                      NaN, value_diff)) %>% 
  group_by(date_time_ID, source) %>% 
  summarise(value_diff_int = sum(value_diff, na.rm = TRUE),
            value_diff_surface = mean(value_diff_surface, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(penetration_depth_diff = value_diff_int / value_diff_surface)

penetration_depth_mean <- penetration_depth %>% 
  group_by(source) %>% 
  summarise(penetration_depth_diff_mean = mean(penetration_depth_diff, na.rm=TRUE),
            penetration_depth_diff_sd = sd(penetration_depth_diff, na.rm=TRUE)) %>% 
  ungroup()

penetration_depth_mean

p_surface <- penetration_depth %>% 
  ggplot(aes(date_time_ID, value_diff_surface, col=source))+
  geom_hline(yintercept = 0)+
  geom_line()+
  geom_point()+
  scale_y_reverse(name = "Change surface value")+
  scale_x_datetime(breaks = "week", date_labels = "%d %b")+
  scale_color_brewer(palette = "Set1", direction = -1)+
  theme(axis.title.x = element_blank(),
        legend.title = element_blank())

p_integrated <- penetration_depth %>% 
  ggplot(aes(date_time_ID, value_diff_int, col=source))+
  geom_hline(yintercept = 0)+
  geom_line()+
  geom_point()+
  scale_y_reverse(name = "Change integrated value")+
  scale_x_datetime(breaks = "week", date_labels = "%d %b")+
  scale_color_brewer(palette = "Set1", direction = -1)+
  theme(axis.title.x = element_blank(),
        legend.title = element_blank())

p_pen_dep <- penetration_depth %>% 
  ggplot(aes(date_time_ID, penetration_depth_diff, col=source))+
  geom_hline(yintercept = 0)+
  geom_hline(data = penetration_depth_mean,
             aes(yintercept = penetration_depth_diff_mean,
                 col=source, linetype = "mean"))+
  geom_line(aes(linetype="cruise"))+
  geom_point()+
  scale_y_reverse(name = "Penetration depth (m)", breaks = seq(0,20,5))+
  scale_x_datetime(breaks = "week", date_labels = "%d %b")+
  scale_color_brewer(palette = "Set1", direction = -1)+
  theme(axis.title.x = element_blank(),
        legend.title = element_blank())


p_surface + p_integrated + p_pen_dep + 
  plot_layout(ncol = 1)

rm(p_surface, p_integrated, p_pen_dep)
rm(penetration_depth, penetration_depth_mean, tm_fm_gt_long)

```


## Remove CO2 profiles

```{r tm_remove_CT_depth, eval=FALSE}

tm_profiles_ID_surface <- tm_profiles_ID %>% 
  filter(dep < surface_dep) %>% 
  group_by(ID) %>% 
  summarise(nCT = mean(nCT)) %>% 
  ungroup() %>% 
  mutate(dep = 3.5)


tm_profiles_ID <- tm_profiles_ID %>% 
  select(-nCT)

tm_profiles_ID <- full_join(tm_profiles_ID, tm_profiles_ID_surface)
rm(tm_profiles_ID_surface)

```


