---
title: "Read raw data"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r packages}
library(tidyverse)
library(data.table)
library(lubridate)
library(DataExplorer)
library(leaflet)
library(readxl)
library(gsubfn)

```

```{r read_parameters, include = FALSE}
parameters <-
  read_rds(here::here("data",
                       "parameters.rds"))
```

```{r ggplot_theme, include = FALSE}
theme_set(theme_bw())
```

# Scope of this script

For each data source:

-   read raw data files
-   harmonize column names
-   clean obvious erroneous data
-   merged into one file
-   write merged file

# Sea-Bird SBE 16 sensor data (ts)

CTD sensor data including recordings from the analog output of auxiliary pH, O~2~, Chla and pCO~2~ sensors were recorded with a measurement frequency of 15 sec. (In addition, pCO~2~ data were also internally recorded on the Contros HydroC instrument with higher temporal resolution and will later be used for further analysis after merging with CTD data.)

## Read regular profiles and transects

```{r read_ts_profile_transect_regular}


files <-
  list.files(path = "data/input/TinaV/Sensor/Profiles_Transects/", pattern = "[.]cnv$")

for (file in files) {
  start_date <-
    data.table(read.delim(
      here::here("data/input/TinaV/Sensor/Profiles_Transects/", file),
      sep = "#",
      nrows = 160
    ))[[78, 1]]
  start_date <- substr(start_date, 15, 34)
  start_date <- mdy_hms(start_date, tz = "UTC")
  
  temp <-
    read.delim(
      here::here("data/input/TinaV/Sensor/Profiles_Transects/", file),
      sep = "",
      skip = 160,
      header = FALSE
    )
  temp <- data.table(temp[, c(2, 3, 4, 5, 6, 7, 9, 11, 13)])
  names(temp) <-
    c("date_time",
      "dep",
      "tem",
      "sal",
      "V_pH",
      "pH",
      "Chl",
      "O2",
      "pCO2_analog")
  temp$start_date <- start_date
  temp$date_time <- temp$date_time + temp$start_date
  
  temp$ID <- substr(file, 1, 6)
  temp$type <- substr(file, 8, 8)
  temp$station <- substr(file, 8, 10)
  
  temp$cast <- "up"
  temp[date_time < mean(temp[dep == max(temp$dep)]$date_time)]$cast <-
    "down"
  
  if (exists("dataset")) {
    dataset <- rbind(dataset, temp)
  }
  
  if (!exists("dataset")) {
    dataset <- temp
  }
  
  rm(start_date)
  rm(temp)
}


ts <- dataset
rm(dataset, file, files)

```

## Read profiles and transects around Östergarnsholm

```{r read_ts_profile_transect_Ostergarnsholm}


files <-
  list.files(path = "data/input/TinaV/Sensor/Ostergarnsholm/", pattern = "[.]cnv$")

for (file in files) {
  start_date <-
    data.table(read.delim(
      here::here("data/input/TinaV/Sensor/Ostergarnsholm/", file),
      sep = "#",
      nrows = 160
    ))[[78, 1]]
  start_date <- substr(start_date, 15, 34)
  start_date <- mdy_hms(start_date, tz = "UTC")
  
  temp <-
    read.delim(
      here::here("data/input/TinaV/Sensor/Ostergarnsholm/", file),
      sep = "",
      skip = 160,
      header = FALSE
    )
  temp <- data.table(temp[, c(2, 3, 4, 5, 6, 7, 9, 11, 13)])
  names(temp) <-
    c("date_time",
      "dep",
      "tem",
      "sal",
      "V_pH",
      "pH",
      "Chl",
      "O2",
      "pCO2_analog")
  temp$start_date <- start_date
  temp$date_time <- temp$date_time + temp$start_date
  
  temp$ID <- substr(file, 1, 6)
  temp$type <- substr(file, 8, 8)
  temp$station <- substr(file, 11, 12)
  
  temp$cast <- "up"
  temp[date_time < mean(temp[dep == max(temp$dep)]$date_time)]$cast <-
    "down"
  
  if (exists("dataset")) {
    dataset <- rbind(dataset, temp)
  }
  
  if (!exists("dataset")) {
    dataset <- temp
  }
  
  rm(start_date)
  rm(temp)
}


ts_OGB <- dataset
rm(dataset, file, files)

ts_OGB <- ts_OGB %>%
  mutate(
    type = if_else(station == "bo", "P", "T"),
    station = if_else(station == "bo", "P14", station),
    station = if_else(station == "in", "T14", station),
    station = if_else(station == "ou", "T15", station)
  )

```

```{r merge_ts_regular_and_Ostergarnsholm}
ts <- bind_rows(ts, ts_OGB) %>%
  arrange(date_time)

rm(ts_OGB)
```

## EDA raw data

```{r ts_EDA_raw, eval=FALSE}

source("code/eda.R")
eda(ts, "ts-raw")
rm(eda)

```

The output of an automated Exploratory Data Analysis (EDA) performed on the raw data with the package `DataExplorer` can be accessed here:

[Link to EDA report of CTD raw data](https://jens-daniel-mueller.github.io/BloomSail/EDA_report_ts-raw.html){target="\"_blank"}

## Clean data set

Sensor recordings were cleaned from obviously erroneous readings, by setting suspicious values to NA.

```{r clean_ts_data}

# running the commented code for plotting
# before and after the cleaning steps
# allows to visualize the removal of errorneous readings

ts <- data.table(ts)

# Profiling data

# temperature

# ts %>%
#   filter(type == "P") %>%
#   ggplot(aes(tem, dep, col=station, linetype = cast))+
#   geom_line()+
#   scale_y_reverse()+
#   geom_vline(xintercept = c(10, 20))+
#   facet_wrap(~ID)

ts[ID == "180723" & station == "P07" & dep < 2 & cast == "up"]$tem <- NA

# salinity

# ts %>%
#   filter(type == "P") %>%
#   ggplot(aes(sal, dep, col=station, linetype = cast))+
#   geom_path()+
#   scale_y_reverse()+
#   facet_wrap(~ID)

ts[sal < 6]$sal <- NA

# pH

# ts %>%
#   filter(type == "P") %>%
#   ggplot(aes(pH, dep, col=station, linetype=cast))+
#   geom_path()+
#   scale_y_reverse()+
#   facet_wrap(~ID)
# 
# ts %>%
#   filter(type == "P") %>%
#   ggplot(aes(V_pH, dep, col=station, linetype=cast))+
#   geom_path()+
#   scale_y_reverse()+
#   facet_wrap(~ID)

ts[pH < 7.5]$V_pH <- NA
ts[pH < 7.5]$pH <- NA

ts[ID == "180709" & station == "P03" & dep < 5 & cast == "down"]$pH <- NA
ts[ID == "180709" & station == "P05" & dep < 10 & cast == "down"]$pH <- NA
ts[ID == "180718" & station == "P10" & dep < 3 & cast == "down"]$pH <- NA
ts[ID == "180815" & station == "P03" & dep < 2 & cast == "down"]$pH <- NA
ts[ID == "180820" & station == "P11" & dep < 15 & cast == "down"]$pH <- NA

ts[ID == "180709" & station == "P03" & dep < 5 & cast == "down"]$V_pH <- NA
ts[ID == "180709" & station == "P05" & dep < 10 & cast == "down"]$V_pH <- NA
ts[ID == "180718" & station == "P10" & dep < 3 & cast == "down"]$V_pH <- NA
ts[ID == "180815" & station == "P03" & dep < 2 & cast == "down"]$V_pH <- NA
ts[ID == "180820" & station == "P11" & dep < 15 & cast == "down"]$V_pH <- NA


# pCO2

# ts %>%
#   filter(type == "P") %>%
#   ggplot(aes(pCO2, dep, col=station, linetype = cast))+
#   geom_path()+
#   scale_y_reverse()+
#   facet_wrap(~ID)

ts[ID == "180616"]$pCO2_analog <- NA

# O2

# ts %>% 
#   filter(type == "P") %>% 
#   ggplot(aes(O2, dep, col=station, linetype = cast))+
#   geom_path()+
#   scale_y_reverse()+
#   facet_wrap(~ID)

# Chlorophyll

# ts %>%
#   filter(type == "P") %>%
#   ggplot(aes(Chl, dep, col=station, linetype = cast))+
#   geom_path()+
#   scale_y_reverse()+
#   facet_wrap(~ID)

ts[Chl > 100]$Chl <- NA


# Surface transect data

# ts %>%
#   filter(type == "T") %>%
#   ggplot(aes(date, dep, col=station))+
#   geom_point()+
#   scale_y_reverse()+
#   facet_wrap(~ID, scales = "free_x")
# 
# ts %>%
#   filter(type == "T") %>%
#   ggplot(aes(date, tem, col=station))+
#   geom_point()+
#   facet_wrap(~ID, scales = "free_x")
# 
# ts %>%
#   filter(type == "T") %>%
#   ggplot(aes(date, sal, col=station))+
#   geom_point()+
#   facet_wrap(~ID, scales = "free_x")
# 
# ts %>%
#   filter(type == "T") %>%
#   ggplot(aes(date, pCO2, col=station))+
#   geom_point()+
#   facet_wrap(~ID, scales = "free_x")
# 
# ts %>%
#   filter(type == "T") %>%
#   ggplot(aes(date, pH, col=station))+
#   geom_point()+
#   facet_wrap(~ID, scales = "free_x")
# 
# ts %>%
#   filter(type == "T") %>%
#   ggplot(aes(date, Chl, col=station))+
#   geom_point()+
#   facet_wrap(~ID, scales = "free_x")

ts[type == "T" & Chl > 10]$Chl <- NA


# ts %>% 
#   filter(type == "T") %>% 
#   ggplot(aes(date, O2, col=station))+
#   geom_point()+
#   facet_wrap(~ID, scales = "free_x")

```

## Write summary file

Relevant columns were selected and renamed, only observations from regular stations (P01-P13) and transects (T01-T13) were selected and summarized data were written to file.

```{r ts_column_subsetting_writing_summary_file}

ts <- ts %>% 
  select(date_time,
         ID,
         type,
         station,
         dep,
         sal,
         tem,
         pCO2_analog)

ts %>% 
  write_csv(here::here("data/intermediate/_summarized_data_files", "ts.csv"))

```

## EDA clean data

```{r ts_EDA_clean, eval=FALSE}

source("code/eda.R")
eda(ts, "ts_clean")

rm(eda)
```

The output of an automated Exploratory Data Analysis (EDA) performed on the cleaned data with the package `DataExplorer` can be accessed here:

[Link to EDA report of CTD clean data](https://jens-daniel-mueller.github.io/BloomSail/EDA_report_ts_clean.html){target="\"_blank"}

## Overview plots

```{r temperature_profiles, fig.cap="Temperature profiles by stations. Color refers to the starting date of each cruise.", fig.asp = 1.3}

ts %>%
  arrange(date_time) %>%
  filter(type == "P",!(station %in% c("PX1", "PX2"))) %>%
  ggplot(aes(tem, dep, col = ymd(ID), group = ID)) +
  geom_path() +
  scale_y_reverse() +
  scale_color_viridis_c(trans = "date", name = "") +
  facet_wrap( ~ station)

```

```{r pCO2_analog_profiles, fig.cap="pCO~2~ (analog signal) profiles by stations. Color refers to the starting date of each cruise.", fig.asp = 1.3}

ts %>%
  arrange(date_time) %>%
  filter(type == "P",!(station %in% c("PX1", "PX2"))) %>%
  ggplot(aes(pCO2_analog, dep, col = ymd(ID), group = ID)) +
  geom_path() +
  scale_y_reverse() +
  scale_color_viridis_c(trans = "date", name = "") +
  facet_wrap( ~ station)

```

# HydroC CO~2~ data (th)

## Read data

Originally, HydroC pCO~2~ data were provided by KM Contros after applying a drift correction to the raw data, which was based on pre- and post-deployment calibration results. Those preliminary data are read-in here. However, some data recorded during testing and configuration of the sensor were later on removed, and the post-processing was repeated based on a cleaned data set. This revised post-processed file is use in the merging script.

```{r read_th_post_processed_by_manufacturer}

# Read Contros corrected data file, based on all recordings
th <-
  read_csv2(here::here("data/input/TinaV/Sensor/HydroC-pCO2/corrected_Contros",
                       "parameter&pCO2s(method 43).txt"),
            col_names = c("date_time", "Zero", "Flush", "p_NDIR",
                          "p_in", "T_control", "T_gas", "%rH_gas",
                          "Signal_raw", "Signal_ref", "T_sensor",
                          "pCO2_corr", "Runtime", "nr.ave")) %>% 
  mutate(date_time = dmy_hms(date_time),
         Flush = as.factor(as.character(Flush)),
         Zero = as.factor(as.character(Zero)))

```

## Deployment identification and subsetting

Individual deployments (periods of observations with less than 30 sec between recordings) were identified and relevant deployment periods were selected. This procedure removes some data recorded during sensor testing and set-up.

```{r identify_select_th_deployment}

# identify individual deployments
th <- th %>%
  arrange(date_time) %>%
  mutate(deployment = cumsum(c(TRUE, diff(date_time) >= 30)))

# write pre-cleaning file for later comparison
th %>%
  select(date_time, pCO2_corr, deployment) %>%
  write_csv(here::here(
    "data/intermediate/_summarized_data_files",
    "th_pre_cleaning.csv"
  ))

# filter relevant deployments
th <- th %>%
  filter(deployment %in% c(2, 6, 9, 14, 17, 21, 23, 27, 31, 33, 34, 35, 37))

```

## Removal of duplicated time stamps

A low number of the recorded HydroC data revealed the exact same time stamp. This was corrected either by filling a corresponding gap before or after the duplicate, or by removing one of the duplicated rows if no such gap existed.

```{r remove_duplicated_timestamp_th}

# add counter for date_time observations
th <- th %>%
  add_count(date_time)

# find triplicated time stamp and select only first observation, and merge

th_no_triple <- th %>%
  filter(n <= 2)

th_triple_clean <- th %>%
  filter(n > 2) %>%
  slice(1)

th <- full_join(th_no_triple, th_triple_clean)

rm(list = setdiff(ls(), c("th", "parameters")))


# find duplicated time stamps and shift first by one second backward, and merge

# th %>%
#   distinct(date_time)

th <- th %>%
  select(-n) %>%
  add_count(date_time)

# unique(th$n)

th_no_duplicated <- th %>%
  filter(n == 1)

th_duplicated <- th %>%
  filter(n == 2)

th_duplicated_first <- th_duplicated %>%
  group_by(date_time) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(date_time = date_time - 1)

th_duplicated_second <- th_duplicated %>%
  group_by(date_time) %>%
  slice(2) %>%
  ungroup()

th_duplicated_clean <-
  full_join(th_duplicated_first, th_duplicated_second) %>%
  arrange(date_time)

th <- full_join(th_no_duplicated, th_duplicated_clean)

# th %>%
#   distinct(date_time)

rm(list = setdiff(ls(), c("th", "parameters")))



# find duplicated time stamps and shift first by two seconds forward, and merge

# th %>%
#   distinct(date_time)

th <- th %>%
  select(-n) %>%
  add_count(date_time)

# unique(th$n)

th_no_duplicated <- th %>%
  filter(n == 1)

th_duplicated <- th %>%
  filter(n == 2)

th_duplicated_first <- th_duplicated %>%
  group_by(date_time) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(date_time = date_time + 2)

th_duplicated_second <- th_duplicated %>%
  group_by(date_time) %>%
  slice(2) %>%
  ungroup()

th_duplicated_clean <-
  full_join(th_duplicated_first, th_duplicated_second) %>%
  arrange(date_time)

th <- full_join(th_no_duplicated, th_duplicated_clean)

# th %>%
#   distinct(date_time)

rm(list = setdiff(ls(), c("th", "parameters")))

# remaining duplicates are observations where other observations with a +/- 1 sec timestamp exist
# for those cases, only the first duplicated observation is selected (similar to triplicate treatment)


# th %>%
#   distinct(date_time)

th <- th %>%
  select(-n) %>%
  add_count(date_time)

# unique(th$n)


th_still_no_duplicated <- th %>%
  filter(n == 1)

th_still_duplicated_first <- th %>%
  filter(n == 2) %>%
  group_by(date_time) %>%
  slice(1)

th <- full_join(th_still_no_duplicated, th_still_duplicated_first)

# th %>%
#   distinct(date_time)

rm(list = setdiff(ls(), c("th", "parameters")))

th <- th %>%
  select(-n)

```

## Flush and Zeroing identification

Flush and zeroing periods of the sensor are identified and assigned with unique IDs.

```{r Identify_Zeroing_Flush_th}

# Zeroing ID labeling

th <- th %>%
  arrange(date_time) %>%
  group_by(Zero) %>%
  mutate(Zero_counter = as.factor(cumsum(c(
    TRUE, diff(date_time) >= 30
  )))) %>%
  ungroup()

# Flush: Identification

th <- th %>%
  mutate(Flush = 0) %>%
  group_by(Zero, Zero_counter) %>%
  mutate(
    start = min(date_time),
    duration = date_time - start,
    Flush = if_else(Zero == 0 &
                      duration < parameters$HC_flush_duration, "1", "0")
  ) %>%
  ungroup()


#  Flush: Identify equilibration and internal gas mixing periods

th <- th %>%
  mutate(mixing = if_else(
    duration < parameters$HC_mixing_duration,
    "mixing",
    "equilibration"
  ))


```

## Deployment plots

```{r th_deployment_plots, eval=FALSE}

pdf(file=here::here("output/Plots/read_in",
    "th_deployments.pdf"), onefile = TRUE, width = 7, height = 4)

for (i in unique(th$deployment)) {
  
  #i <- unique(th$deployment)[3]
  
  sub <-  th %>%
      filter(deployment == i)
  start_date <- min(sub$date_time)
  
  print(
    sub %>% 
      ggplot(aes(date_time, pCO2_corr, col=Zero_counter))+
      geom_line()+
      labs(title = paste("Deployment: ",i, "| Start time: ", start_date))
)
  
}

dev.off()

rm(sub, start_date, i)

```

A pdf with pCO~2~ timeseries plots of all individual deployments can be found here:

[Link to pCO~2~ timeseries plots](https://github.com/jens-daniel-mueller/BloomSail/tree/master/output/Plots/read_in/th_deployments.pdf){target="_blank"}

```{r th_EDA, eval=FALSE}

source("code/eda.R")
eda(th, "th")
rm(eda)

```

The output of an automated Exploratory Data Analysis (EDA) performed with the package `DataExplorer` can be accessed here:

[Link to EDA report of HydroC pCO~2~ data data](https://jens-daniel-mueller.github.io/BloomSail/EDA_report_th.html){target="\"_blank"}

## Write summary file

Summarized clean pCO~2~ data were written to file.

```{r write_th_summary_file}

th %>%
  select(date_time,
         Zero,
         Flush,
         pCO2_corr,
         deployment,
         Zero_counter,
         duration,
         mixing) %>%
  write_csv(here::here("data/intermediate/_summarized_data_files",
                       "th.csv"))

rm(th)

```

# Bottle data CO~2~ (tb)

Discrete samples were collected with a Niskin bottle and analyzed for CT and AT at IOW's CO~2~ lab.

## Read data

```{r read_tb_CO2_data}

# Read CO2 system bottle data
tb <-
  read_csv(
    here::here(
      "data/input/TinaV/Bottle/Tracegases",
      "BloomSail_bottle_CO2_all.csv"
    ),
    col_types = list("c", "c", "n", "n", "n", "n", "n")
  )

# select and rename relevant columns
tb <- tb %>%
  select(
    ID = transect.ID,
    station = label,
    dep = Dep,
    sal = Sal,
    CT,
    AT
  )


```

## Write summary file

```{r write_tb_CO2_data}

tb %>% write_csv(here::here("data/intermediate/_summarized_data_files",
                            "tb.csv"))
rm(tb)

```

# Bottle data plankton (tp)

Discrete samples were collected with a Niskin bottle and analysed for phytoplankton composition and biomass at IOW's phytoplankton lab (Norbert Wasmund).

## Read data

```{r read_tp_phytoplankton_data}

tp <- read_csv(
  here::here(
    "data/input/TinaV/Bottle/Phytoplankton",
    "181205_BloomSail_Plankton_counts.csv"
  )
)

# delete colomns that contain counts, not calculated biomass
tp <- tp[, -seq(4, 21, 1)]

# assign new column names
# for species: 
# nr = size class, 
# HV = Heterocyst per Volume, 
# Hl = Heterocyst per length, 
# t = total

names(tp) <-
  c(
    "date",
    "station",
    "dep",
    "Aphanizomenon.1",
    "Aphanizomenon.2",
    "Aphanizomenon.3",
    "Aphanizomenon.t",
    "Aphanizomenon.HV",
    "Aphanizomenon.Hl",
    "Dolichospermum.1",
    "Dolichospermum.2",
    "Dolichospermum.3",
    "Dolichospermum.4",
    "Dolichospermum.t",
    "Dolichospermum.HV",
    "Dolichospermum.Hl",
    "Nodularia.1",
    "Nodularia.2",
    "Nodularia.3",
    "Nodularia.t",
    "Nodularia.HV",
    "Nodularia.Hl",
    "Nodulariadead.1",
    "Nodulariadead.2",
    "Nodulariadead.3",
    "Nodulariadead.t",
    "total.t"
  )


# change format of data table and separate into 2 columns for species and class

tp <-
  gather(tp, para, value, Aphanizomenon.1:total.t, factor_key = TRUE)

tp <- separate(tp, col = para, into = c("Species", "class"))

# change class of columns
tp <- tp %>%
  mutate(ID = date,
         date = ymd(date))

```

## Write summary file

```{r write_tp_phytoplankton_data}

tp %>% write_csv(here::here("data/intermediate/_summarized_data_files",
                            "tp.csv"))
rm(tp)

```

# GPS track (tt)

GPS track data were recorded with a Samsung Galaxy tablet.

## Read data

```{r read_tt_data}

files <-
  list.files(path = "data/input/TinaV/Track/GPS_Logger_Track/", pattern = "[.]txt$")

for (file in files) {
  # if the merged dataset does exist, append to it
  if (exists("dataset")) {
    temp <-
      data.table(read.delim(
        here::here("data/input/TinaV/Track/GPS_Logger_Track", file),
        sep = ","
      )[, c(2, 3, 4)])
    names(temp) <- c("date_time", "lat", "lon")
    temp$date_time <- ymd_hms(temp$date, tz = "UTC")
    
    dataset <- rbind(dataset, temp)
    rm(temp)
  }
  
  # if the merged dataset doesn't exist, create it
  if (!exists("dataset")) {
    dataset <-
      data.table(read.delim(
        here::here("data/input/TinaV/Track/GPS_Logger_Track", file),
        sep = ","
      )[, c(2, 3, 4)])
    names(dataset) <- c("date_time", "lat", "lon")
    dataset$date_time <- ymd_hms(dataset$date_time, tz = "UTC")
    
  }
}

tt <- dataset
rm(dataset, file, files)

```

## Write summary file

```{r write_tt_data}

tt %>%
  write_csv(here::here("data/intermediate/_summarized_data_files",
                       "tt.csv"))

rm(tt)

```

# Atmospheric observations

Atmospheric data were recorded at the ICOS station on Östergarnsholm.

## Read data

```{r read_wind_Ostergarn_tower}

og <-
  read_delim(
    here::here(
      "data/input/Ostergarnsholm/Tower",
      "Oes_Jens_atm_water_June_to_August_2018.csv"
    ),
    delim = ";"
  )

og <- og %>%
  mutate(date_time = ymd_hms(paste(
    paste(year, month, day, sep = "/"),
    paste(hour, min, sec, sep = ":")
  ))) %>%
  select(
    "date_time",
    "CO2 12m [ppm]",
    "w_c [ppm m/s]",
    "WS 12m [m/s]",
    "WD 12m [degrees]",
    "T 12m [degrees C]",
    "RIS [W/m^2]"
  )

# conversion from GMT+1 to UTC
og <- og %>%
  mutate(date_time = date_time - 60 ^ 2)

og <- og %>%
  select(date_time, pCO2_atm = "CO2 12m [ppm]", wind = "WS 12m [m/s]")

```

## Write summary file

```{r write_og_data}

og %>%
  write_csv(here::here("data/intermediate/_summarized_data_files",
                       "og.csv"))

rm(og)

```

# SOOP Finnmaid

Here, we read in pCO~2~ and SST data recorded on SOOP Finnmaid in June-August 2018.

## Read data

```{r read_fm_data}

# LI-COR data

files <-
  list.files(path = "data/input/Finnmaid_2018", pattern = "[.]xls$")

for (file in files) {
  temp <- read_excel(here::here("data/input/Finnmaid_2018", file))
  temp <- temp[c(1, 2, 3, 12, 7, 4, 15, 8, 5, 17)]
  names(temp) <-
    c("date_time",
      "lon",
      "lat",
      "pCO2",
      "sal",
      "tem",
      "cO2",
      "patm",
      "Teq",
      "xCO2")
  temp <- temp[-c(1), ]
  temp$date_time <-
    as.POSIXct(as.numeric(temp$date_time) * 60 * 60 * 24,
               origin = "1899-12-30",
               tz = "GMT")
  temp$lon <- as.numeric(as.character(temp$lon))
  temp$lat <- as.numeric(as.character(temp$lat))
  temp$pCO2 <- as.numeric(as.character(temp$pCO2))
  temp$sal <- as.numeric(as.character(temp$sal))
  temp$tem <- as.numeric(as.character(temp$tem))
  temp$cO2 <- as.numeric(as.character(temp$cO2))
  temp$patm <- as.numeric(as.character(temp$patm))
  temp$Teq <- as.numeric(as.character(temp$Teq))
  temp$xCO2 <- as.numeric(as.character(temp$xCO2))
  temp <- data.table(temp)
  
  temp$route <-
    strapplyc(as.character(file), ".*(.).xls*", simplify = TRUE)
  temp$ID <- substr(as.character(file), 3, 10)
  
  if (exists("dataset")) {
    dataset <- rbind(dataset, temp)
  } else{
    dataset <- temp
  }
  
}


rm(temp, files, file)
dataset <- dataset[pCO2 != 0]


# Los Gatos Research (LGR) data
# Please note that the LGR data were corrected manually before
# The correction procedure is outlined in the Appendix of the ms

files <-
  list.files(path = "data/input/Finnmaid_2018/LGR", pattern = "[.]xls$")

for (file in files) {
  temp <- read_excel(here::here("data/input/Finnmaid_2018/LGR", file))
  temp <- temp[c(2, 3, 4, 8, 6, 5, 14, 7, 15, 9)]
  names(temp) <-
    c("date_time",
      "lon",
      "lat",
      "pCO2",
      "sal",
      "tem",
      "cO2",
      "patm",
      "Teq",
      "xCO2")
  temp <- temp[-c(1), ]
  temp$date_time <- dmy_hms(temp$date_time)
  temp <- data.table(temp)
  
  temp$route <- substr(as.character(file), 12, 12)
  temp$ID <- substr(as.character(file), 3, 10)
  
  if (exists("dataset.LGR")) {
    dataset.LGR <- rbind(dataset.LGR, temp)
  } else{
    dataset.LGR <- temp
  }
  
}

rm(temp, files, file)


```

```{r convert_O2_Los_Gatos_fm, eval=FALSE}

# This code can be used to convert O2 units
# but is not applied, because O2 data are not used in this study

source(here::here("code", "O2stoO2c.R"))

dataset.LGR <- dataset.LGR %>%
  filter() %>%
  mutate(cO2 = O2stoO2c(
    O2sat = cO2,
    T = tem,
    S = sal,
    P = 3 / 10,
    p_atm = 1013.5
  ))

rm(O2stoO2c, pH2Osat, sca_T, Scorr, TCorr, R, Vm)

```

```{r merge_Los_Gatos_LICOR_data_files}

dataset$sensor <- "LICOR"
dataset.LGR$sensor <- "LosGatos"

fm <- bind_rows(dataset, dataset.LGR)

rm(dataset, dataset.LGR)

```

## Write summary file

```{r write_fm}

fm %>% 
  write_csv(here::here("data/intermediate/_summarized_data_files",
                       "fm.csv"))
rm(fm)

```

# Interactive map

```{r map, fig.cap="Overview map of SOOP Finnmaid and SV Tina V tracks. Please note that only a subet of the recorded track data is plotted."}


fm <-
  read_csv(here::here("data/intermediate/_summarized_data_files",
                      "fm.csv"))

fm_sub <- fm %>%
  arrange(date_time) %>%
  slice(which(row_number() %% 20 == 1))

tt <-
  read_csv(here::here("data/intermediate/_summarized_data_files", "tt.csv"))

tt_sub <- tt %>%
  slice(which(row_number() %% 20 == 1))

rm(tt, fm)

leaflet() %>%
  setView(lng = 20, lat = 57.3, zoom = 8) %>%
  addLayersControl(
    baseGroups = c("Ocean Basemap",
                   "Satellite"),
    overlayGroups = c("BloomSail", "Finnmaid"),
    options = layersControlOptions(collapsed = FALSE),
    position = 'topright'
  ) %>%
  addProviderTiles("Esri.WorldImagery", group = "Satellite") %>%
  addProviderTiles(providers$Esri.OceanBasemap, group = "Ocean Basemap") %>%
  addScaleBar(position = 'topright') %>%
  addMeasure(
    primaryLengthUnit = "kilometers",
    secondaryLengthUnit = 'miles',
    primaryAreaUnit = "sqmeters",
    secondaryAreaUnit = "acres",
    position = 'topleft'
  ) %>%
  addCircles(data = fm_sub,
             ~ lon,
             ~ lat,
             color = "white",
             group = "Finnmaid") %>%
  addPolylines(data = tt_sub,
               ~ lon,
               ~ lat,
               color = "red",
               group = "BloomSail")


rm(fm_sub, tt_sub)

```
