---
title: "Reconstruction of NCP based on surface pCO2 and CTD data"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---


```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r packages}
library(tidyverse)
library(seacarb)
```

# Sensor data

1m gridded, downcast profiles were used.

```{r read_Tina-V_Sensor_data}

df <-
  read_csv(here::here("Data/_merged_data_files", "BloomSail_CTD_HydroC_CT_cumulative_profiles.csv"))

df <- df %>% 
  select(1:5) %>% 
  pivot_wider(names_from = "parameter", values_from = "value") %>% 
  mutate(CT = if_else(dep==3.5, CT, NaN))

```

# C~T~ vs temperature

As primary production (negative changes in C~T~) and increase in seawater temperature have a common driver (light), the relation between both changes was investigated.

```{r CT_vs_temperature}

df_CT <- df %>%
  drop_na() %>% 
  arrange(date_time_ID) %>%
  mutate(dCT = CT - lag(CT))

df <- df %>%
  arrange(date_time_ID) %>%
  mutate(dtem = tem - lag(tem))

df <- full_join(df, df_CT)
rm(df_CT)

df %>% 
  ggplot()+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  geom_smooth(aes(dtem, dCT), method = "lm", se=FALSE)+
  geom_point(aes(dtem, dCT), shape=21)+
  scale_fill_viridis_d()+
  guides(fill = guide_legend(override.aes=list(shape=21)))

CT_tem %>% 
  ggplot(aes(dtem, dCT))+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  geom_smooth(method = "lm", se=FALSE)+
  geom_point(aes(fill=ID), shape = 21)+
  scale_fill_viridis_d()+
  guides(fill = guide_legend(override.aes=list(shape=21)))+
  facet_wrap(~ID)

```




## Profiles of incremental changes

Changes of seawater parameters at each depth are calculated from one cruise day to the next and divided by the number of days inbetween.

```{r incremental_changes_profiles}

mean_profiles_long <- mean_profiles_long %>%
    group_by(parameter, dep) %>%
    arrange(date_time_ID) %>%
    mutate(diff_value = value     - lag(value, default = first(value)),
           diff_time  = as.numeric(date_time_ID - lag(date_time_ID)),
           diff_value_daily = diff_value / diff_time,
           cum_value = cumsum(diff_value)) %>% 
  ungroup()


mean_profiles_long %>% 
  arrange(dep) %>% 
  ggplot(aes(diff_value_daily, dep, col=ID))+
  geom_vline(xintercept = 0)+
  geom_point()+
  geom_path()+
  scale_y_reverse()+
  scale_color_viridis_d()+
  facet_wrap(~parameter, scales = "free_x")+
  labs(x="Change of value inbetween cruises per day")
  
```
